### 2019-04-15 presentation/pres.tex

\documentclass[slidestop,compress,red,mathserif]{beamer}

\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
\usepackage{amsmath,amssymb}
\usepackage{comment}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{marvosym}
\usepackage{natbib}
\usepackage{ulem}
\usepackage{wasysym}

\setbeamersize{text margin left=6mm, text margin right=2mm}

\mode<presentation>
{
%\usetheme{Warsaw}
%\usetheme{Rochester}
%\usetheme{Madrid}
%\usetheme{Pittsburgh}
%\usetheme{Antibes}
%\usetheme{Montpellier}
%\usetheme{Berkeley}
%\usetheme{PaloAlto}
%\usetheme{Goettingen}
%\usetheme{Marburg}
%\usetheme{Hannover}
%\usetheme{Berlin}
\usetheme{Ilmenau}
%\usetheme{Dresden}
%\usetheme{Darmstadt}
%\usetheme{Frankfurt}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Copenhagen}
%\usetheme{Malmoe}
\setbeamercovered{transparent}
}

% Choose color scheme
%
\usecolortheme{default}
%\usecolortheme{sidebartab}
%\usecolortheme{albatross}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{seagull}
%\usecolortheme{lily}
%\usecolortheme{orchid}

\beamertemplateballitem
%\setbeamertemplate{footline}{\insertframenumber/\inserttotalframenumber}

\defbeamertemplate*{footline}{shadow theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=2.\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm plus1fil,rightskip=.3cm]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertframenumber\,/\,\inserttotalframenumber\hfill%\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
    \usebeamerfont{title in head/foot}%\insertshorttitle%
  \end{beamercolorbox}}%
  \vskip0pt%
}

% number sets
\def\N{\hbox{I \kern-.5em N}}
\def\R{\hbox{I \kern-.5em R}}
\def\Q{\hbox{\hspace*{.25em}{\sf I} \kern-.83em Q}}
\def\T{\rm T}  %%%%  \rm pour roman (ecriture normale)
\def\Z{\rm Z}
\def\ZZ{\hbox{\sf  Z \kern-.8em Z}}
\def \Zplus        {\ZZ^{+}}
\def \bigm {\textsc{penal}}
\def \dst {\textsc{dst}}
\def \src {\textsc{src}}
\def \zlp {z^{\star}_{\textsc{lp}}}
\def \zilp {{\tilde z}_{\textsc{ilp}}}
\def \l    {\textnormal{\textsc{l}}}
\def \Gp   {G_{\textsc{p}}}
\def \Gl   {G_{\l}}
\def \Vp   {V_{\textsc{p}}}
\def \Vl   {V_{\l}}
\def \Ep   {E_{\textsc{p}}}
\def \El   {E_{\l}}
\def \cost {\textsc{cost}}
\def \srlg {\textsc{srlg}}

% number sets
\def \best {\textsc{best}}
\def  \N            {\hbox{I \kern-.5em N}}
\def \RR {\mathbb{R}}
\def \Rplus {\RR^+}
\def  \Z            {\rm Z}
\def  \Zplus        {\Z^+}
\def  \Zplusn       {\Zplus_n}
\def  \ZZ           {\hbox{\sf  Z \kern-.8em Z}}
\def \Z {\mathbb{Z}}

\logo{%
    \includegraphics[width=2.cm,keepaspectratio]{Figures/ConULogo_CMYK.eps}~%
}

\title[My Title]{Accident prediction in time and space}
\author[My~Name]{
\textbf{Timothée Guédon, Antoine Hébert}}
\institute[My Institute/company]
{Concordia University \\ Computer Science and Software Engineering Department}
\date[]{April 15, 2019}

\begin{document}

\begin{frame} % Cover slide
\titlepage
\end{frame}

% layout of the presentation
\AtBeginSubsection[]
{
  \begin{frame}<beamer>
    \frametitle{Layout}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection]
  \end{frame}
}

%-------------------------------------------------
\section{Introduction}
%-------------------------------------------------

% -------------------------------------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Road accident prediction}

\begin{scriptsize}

\begin{itemize}
\item[] \textbf{Context}:
  \begin{itemize}
    \item Road accidents implies millions of people deaths and injuries every year.
		\item Human and economic impact on society.
    \item Master’s thesis of Antoine Hébert in a different context.
  \end{itemize}

\item[] \textbf{Our goals}:
  \begin{itemize}
    \item Dataset analysis.
  	\item Finding key insights in accident prediction.
  	\item Identifying issues related to class imbalance and geo-spatial data.
  \end{itemize}
\end{itemize}

\end{scriptsize}

\end{frame}

% -------------------------------------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Datasets}

\begin{footnotesize}
We used open datasets from Open Canada.

\begin{itemize}
\item[] \textbf{National Collision Database (NCDB)}:
  \begin{itemize}
  	\item All vehicle collisions reported by the police from 2012 to 2017 in Montreal.
  	\item Date and time, severity, number of death and injury, number of vehicles, localization, speed limit etc.
  \end{itemize}
\item[] \textbf{National Road Network}:
  \begin{itemize}
    \item Shape of the road, how roads are
connected together, the mail addresses on the street, the number of lanes in each direction and the
type of pavement.
  \end{itemize}
\item[] \textbf{Climate data - Hourly Data Report}:
  \begin{itemize}
    \item For each position and hour, the dataset provides the temperature, the humidity, the wind speed and direction, the visibility, the atmospheric pressure etc.
  \end{itemize}
\end{itemize}

\end{footnotesize}

\end{frame}

% -------------------------------------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Tree-based algorithms}
We chose to focus on tree-based algorithms because it has proven its efficiency on accident prediction.
\begin{itemize}
\item[] \textbf{Related work on accident prediction}:
  \begin{itemize}
    \item Extensive use of decision trees in several form. (Athanasios Theofilatos, A. 2017, Joaquín Abellán, 2013, Lei Lin 2014, Li-Yen Chang 2005)
    \item Recently, deep learning algorithms (LSTM and CNN architectures for example) (Zhuoning Yuan, 2018, Quanjun Chen 2016)
  \end{itemize}
\item[] \textbf{Chosen algorithms}:
  \begin{itemize}
  	\item Random Forest
    \item Balanced Random Forest
    \item Gradient Boosted Trees (GBT) - XGBoost implementation
  \end{itemize}
\end{itemize}
\end{frame}

%-------------------------------------------------
\section{Methods}
%-------------------------------------------------

\begin{frame}
	\frametitle{Generating the negative samples}
	\begin{itemize}
    \item Every time an accident did not happen is considered a negative sample.
		\item A sample from the cartesian product between all dates (between 01/01/2012 and 31/12/2017) and all roads.
		\item Generation of 5 million negative samples for about 200 000 positive samples.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Some feature engineering}
	\begin{itemize}
		\item Computing the distances between GPS coordinates (taking into account earth curvature)
		\item Making the dates and times cyclic
		\item Creating missing points on the roads
    \item Interpolating weather statistics by averaging data from several stations
    \item etc...
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Area under PR curve}
	\begin{itemize}
    \item ROC Curve : false positive rate VS true positive rate
    \begin{equation*}
			FPR = \frac{FP}{FP+TN},  TPR = \frac{TP}{TP+FN}
		\end{equation*}
		\item PR Curve : true positive rate VS precision
    \begin{equation*}
			TPR = \frac{TP}{TP+FN},  Precision = \frac{TP}{TP+FP}
		\end{equation*}
    \item Data imbalance imply decrease of TP and increase of FN
    \item Therefore, area under PR exhibits larger difference than are under ROC
	\end{itemize}
\end{frame}

%-------------------------------------------------
\section{Balanced Random Forests}
%-------------------------------------------------

\begin{frame}
\frametitle{Remainder: Random Forests}
\begin{itemize}
  \item[] \textbf{Remainder}
    \begin{itemize}
    	\item Random Forests (\cite{Breiman2001}) are combinations of fully-grown decision trees.
    	\item Each tree is trained on a random sample of the data and a random sample of the features.
    \end{itemize}

  \item[] \textbf{Weighted Random Forests (WRF)}:
    \begin{itemize}
      \item A solution to imbalance.
      \item Add weights to the classes when growing the tree (in the Gini coefficient).
      \item Add weights to the classes at the voting time.
    \end{itemize}

\end{itemize}
~~~~~~~~~~~~~~~
\end{frame}

\begin{frame}
\frametitle{Solution to imbalance}
\begin{itemize}
  \item[] \textbf{Balanced Random Forests (BRF)}:
    \begin{itemize}
      \item Another solution to imbalance.
      \item Combine undersampling and ensemble method versus loss of information.
      \item For each tree : Randomly draw sample of same size from majority (with replacement) and minority class.
      \item The rest of the algorithm is identical to Random Forests
    \end{itemize}

  \item[] \textbf{BRF versus WRF}:
    \begin{itemize}
      \item No clear winner ! But...
      \item WRF - More vulnerable to noise on minority class
      \item BRF - More computationally efficient because use less data (subsampling)
      \item BRF - Easier to implement
    \end{itemize}

\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Implementation of BRF in Spark}
\begin{itemize}
  \item[] \textbf{Existing implementation of RF in Spark}
    \begin{itemize}
    	\item To manage the subsampling of the dataset, Spark's scala implementation use the Poisson distribution.
      \item The Poisson distribution is used to compute the probability of an event to occur at a given time.
    	\item To know how many times a given sample will be in the subsample of a given tree (can be 0 times), we use "distribution sampling" of 1 element.
      \item (Distribution sampling: Get a random sample following a given distribution)
    \end{itemize}
  \item[] \textbf{Our modifications}:
    \begin{itemize}
      \item Instead of creating one Poisson distribution (with the "p" parameter and a seed) we create two Poisson distributions:
    	\item[]
      \begin{itemize}
      	\item One for the distribution of the "positive" samples,
        \item Another for the distribution of the "negative" samples.
      \end{itemize}
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Experimenting with our BRF implementation}
\begin{itemize}
  \item Area Under PR : 99,3\% for WRF VS 99,7\% for BRF
  \item F1 score : 99\% for WRF VS 98\% for BRF (which proves that F1 score can be misleading)

\end{itemize}

\centering
\includegraphics[height=3.5cm, keepaspectratio]{Figures/BRF.png}
\includegraphics[height=3.5cm, keepaspectratio]{Figures/WRF.png}

\end{frame}

%-------------------------------------------------
\section{Big Data related issues}
%-------------------------------------------------

\begin{frame}
	\frametitle{Dataset generation}
	\begin{itemize}
		\item Using slurm-based (scheduler) cluster to increase the dataset generation's speed.
    \item Tricks to increase the generation speed and reduce the memory consumption.
	\end{itemize}
\centering
\includegraphics[height=3.5cm, keepaspectratio]{Figures/timeVSsamples_generatepositivesamples.png}
\includegraphics[height=3.5cm, keepaspectratio]{Figures/timeVSsamples_generatepositivesamples_after.png}
\end{frame}

\begin{frame}
	\frametitle{Spark VS Dask}
	\begin{itemize}
    \item[] \textbf{Advantages of using Dask over Spark}
      \begin{itemize}
      	\item Mimic the pandas dataframe API
        \item Lighter than Spark
      	\item Designed to be more flexible in terms of applications and algorithms
      \end{itemize}
    \item[] \textbf{Why we switched for Spark}
      \begin{itemize}
        \item We found ourselves writing a lot more code using Dask than using Spark
        \item We found Spark easier to use for processing the dataset in batch
        \item Spark's map-reduce and shuffling operations are very handy
        \item Spark ML pipelines allow to create pipelines of preprocessing easily
      \end{itemize}
	\end{itemize}
\end{frame}

%-------------------------------------------------
\section{Results}
%-------------------------------------------------

\begin{frame}
	\frametitle{Comparison between the algorithms}
	\begin{itemize}
		\item Random forest
		\item Balanced random forest
		\item XGBoost
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Key features identified}
	\begin{itemize}
		\item The most important features seem to be the time and location of the vehicle, followed by some road's features like the type of road (highway for example) and the street length.
    \item Finally, comes wome weather features like the visibility, the wind and the humidity.
	\end{itemize}
\centering
\includegraphics[height=4cm, keepaspectratio]{Figures/features.png}
\end{frame}

%-------------------------------------------------
\section{Conclusion}
%-------------------------------------------------

\begin{frame}
	\frametitle{Conclusion}
	\begin{itemize}
    \item We found some key features that can explain the occurence of accidents.
    \item We identified some issues related to class imbalance and big data analytics.
    \item Future work would include tuning the XGBoost classifier, testing lightGBM classifier, assing some datasets like a road works dataset.
    \item The code can be found on Github at https://github.com/GTimothee/accident-prediction-montreal
	\end{itemize}
\end{frame}

% -------------------------------------------------------------------------------------------------------------------------
% -------------------------------------------------------------------------------------------------------------------------
%% -------------------------------------------------------------------------------------------------------------------------

\begin{frame}[allowframebreaks]
\frametitle{References}

{\scriptsize

\bibliographystyle{plainnat}
\bibliography{Biblio}

}

\end{frame}

\end{document}

### 2019-04-15 main_xgboost.py
Add workdir to main_xgboost.py

@@ -8,6 +8,7 @@ from utils import init_spark
 from preprocess import get_positive_samples, \
                        get_negative_samples, \
                        get_dataset_df
+from workdir import workdir
 
 
 from pyspark.sql import SparkSession
@@ -24,7 +25,7 @@ from pyspark.ml.tuning import ParamGridBuilder, CrossValidator
 
 def save_model():
     workdir = "./"
-    path = "data/xgboost.model"
+    path = workdir + "data/xgboost.model"
     if os.path.isdir(path):
         shutil.rmtree(path)
     model.save(path)
@@ -32,7 +33,7 @@ def save_model():
 
 
 def grid_search(spark):
-    sample_ratio = 0.1
+    sample_ratio = 0.05
     neg_samples = get_negative_samples(spark).sample(sample_ratio).na.fill(0)
     pos_samples = get_positive_samples(spark).sample(sample_ratio).na.fill(0)
     df = get_dataset_df(spark, pos_samples, neg_samples).na.fill(0)
@@ -72,6 +73,6 @@ def grid_search(spark):
 
 spark = init_spark()
 #os.environ['PYSPARK_SUBMIT_ARGS'] = '--jars data/xgboost4j-spark-0.72.jar,data/xgboost4j-0.72.jar pyspark-shell'
-spark.sparkContext.addPyFile("data/sparkxgb.zip")
+spark.sparkContext.addPyFile(workdir + "data/sparkxgb.zip")
 from sparkxgb import XGBoostEstimator
 grid_search(spark)

### 2019-04-15 accidents_montreal.py
Change cache file names just in case

@@ -45,7 +45,7 @@ def fetch_accidents_montreal():
 
 
 def read_accidents_montreal_df(spark, use_cache=True):
-    cache = workdir + 'data/accidents.parquet'
+    cache = workdir + 'data/accidents_montreal.parquet'
     if (os.path.isdir(cache) or os.path.isfile(cache)) and use_cache:
         print('Skip extraction of accidents montreal dataframe:'
               ' already done, reading from file')

### 2019-04-15 main_gen_match.py

@@ -17,7 +17,7 @@ from accidents_montreal import get_accident_df
 
 
 def generate_match_accident_road_of_one_month(year, month):
-    filepath = workdir + f'data/match_accident_road_{year}-{month}.parquet'
+    filepath = workdir + f'data/match_accident-road_{year}-{month}.parquet'
     if isdir(filepath):  # Skip if already done
         return
     print(f'Generating {year}-{month}')
@@ -57,12 +57,12 @@ def match_accident_road_samples_reader():
     for year in range(2012, 2018):
         for month in range(1, 13):
             filepath = workdir \
-                + f'data/match_accident_road_{year}-{month}' \
+                + f'data/match_accident-road_{year}-{month}' \
                 + '.parquet'
             yield spark.read.parquet(filepath)
 
 
 full_matching_accident_road = reduce(DataFrame.union,
                                      match_accident_road_samples_reader())
-filepath = workdir + 'data/match_accident_road.parquet'
+filepath = workdir + 'data/match_accident-road.parquet'
 full_matching_accident_road.write.parquet(filepath)

### 2019-04-15 preprocess.py

@@ -38,7 +38,7 @@ def preprocess_accidents(accidents_df):
 
 
 def match_accidents_with_roads(spark, road_df, accident_df, use_cache=True):
-    cache_path = workdir + 'data/match_accident_road.parquet'
+    cache_path = workdir + 'data/match_accident-road.parquet'
     if isdir(cache_path) and use_cache:
         print('Reading accident-road matches from cache...')
         return spark.read.parquet(cache_path)

### 2019-04-15 main_random_forest.py
Fix main_random_forest.py

@@ -13,17 +13,18 @@ from workdir import workdir
 
 spark = init_spark()
 
-neg_samples = get_negative_samples(spark).sample(0.1)
-pos_samples = get_positive_samples(spark).sample(0.1)
+neg_samples = get_negative_samples(spark).sample(0.01)
+pos_samples = get_positive_samples(spark).sample(0.01)
 df = get_dataset_df(spark, pos_samples, neg_samples)
 df_sample = df.filter(col('date') > datetime.fromisoformat('2017-01-01'))
 (train_set, test_set) = df_sample.randomSplit([0.7, 0.3])
 
 model = random_forest_tuning(train_set)
 
-with open(workdir + 'data/random_forest_tuning_results4.txt', 'w') as file:
-    for model, result in zip(model.subModels, model.validationMetrics):
+with open(workdir + 'data/random_forest_tuning_results5.txt', 'w') as file:
+    for models, result in zip(model.subModels, model.avgMetrics):
         file.write('==================================\n')
+        model = models[0]
         for stage in model.stages:
             params = stage.extractParamMap()
             for k in params:
@@ -32,8 +33,8 @@ with open(workdir + 'data/random_forest_tuning_results4.txt', 'w') as file:
 
 predictions = model.transform(test_set)
 area_under_PR, f1_score = evaluate_binary_classifier(predictions)
-with open(workdir + 'data/random_forest_tuning_perf_4.txt', 'w') as file:
+with open(workdir + 'data/random_forest_tuning_perf_5.txt', 'w') as file:
     file.write(f"Area Under PR = {area_under_PR}\nF1 score = {f1_score}")
 
 graph = compute_precision_recall_graph(predictions, 20)
-graph.to_csv(workdir + 'data/precision_recall_graph_rf_4.csv')
+graph.to_csv(workdir + 'data/precision_recall_graph_rf_5.csv')

### 2019-04-15 main_xgboost.py
Make main_xgboost.py outputs results

@@ -64,15 +64,28 @@ def grid_search(spark):
                         .asInstanceOf[PipelineModel]
                         .stages(2)
                         .asInstanceOf[XGBoostClassificationModel])
-    bestModel.extractParamMap()
+
+    with open(workdir + 'data/xgboost_tuning_results_1.txt', 'w') as file:
+        for models, result in zip(model.subModels, model.avgMetrics):
+            file.write('==================================\n')
+            model = models[0]
+            for stage in model.stages:
+                params = stage.extractParamMap()
+                for k in params:
+                    file.write(f'{k.name}: {params[k]}\n')
+            file.write(f"Area under PR: {result}\n")
+
     prediction = bestModel.transform(testDF)
     prediction = prediction.withColumn("rawPrediction", prediction['probabilities'])
     area_under_PR, f1_score = evaluate_binary_classifier(prediction)
+
+    with open(workdir + 'data/xgboost_tuning_perf_1.txt', 'w') as file:
+        file.write(f"Area Under PR = {area_under_PR}\nF1 score = {f1_score}")
+
     return
 
 
 spark = init_spark()
-#os.environ['PYSPARK_SUBMIT_ARGS'] = '--jars data/xgboost4j-spark-0.72.jar,data/xgboost4j-0.72.jar pyspark-shell'
 spark.sparkContext.addPyFile(workdir + "data/sparkxgb.zip")
 from sparkxgb import XGBoostEstimator
 grid_search(spark)

### 2019-04-15 random_forest.py
Fix random_forest.py and change param grid

@@ -1,7 +1,7 @@
 from pyspark.ml.classification import RandomForestClassifier
 from pyspark.ml.evaluation import BinaryClassificationEvaluator
 from pyspark.ml.evaluation import MulticlassClassificationEvaluator
-from pyspark.ml.tuning import TrainValidationSplit, ParamGridBuilder
+from pyspark.ml.tuning import CrossValidator, ParamGridBuilder
 from pyspark.ml import Pipeline
 from pyspark.sql.types import DoubleType
 from pyspark.sql.functions import udf, col
@@ -19,23 +19,22 @@ def random_forest_tuning(train_samples):
     pipeline = Pipeline().setStages([ru, rf])
     paramGrid = \
         (ParamGridBuilder()
-         .addGrid(rf.numTrees, [100, 110, 115, 120, 130])
+         .addGrid(rf.numTrees, [100, 115, 130])
          .addGrid(rf.featureSubsetStrategy, ['sqrt'])
          .addGrid(rf.impurity, ['entropy'])
          .addGrid(rf.maxDepth, [30])
          .addGrid(rf.minInstancesPerNode, [1])
-         .addGrid(rf.subsamplingRate, [0.8])
-         .addGrid(ru.targetImbalanceRatio, [1.0])
+         .addGrid(rf.subsamplingRate, [0.8, 0.6])
+         .addGrid(ru.targetImbalanceRatio, [1.0, 1.5, 1.25])
          .build())
     pr_evaluator = \
         BinaryClassificationEvaluator(labelCol="label",
                                       rawPredictionCol="rawPrediction",
                                       metricName="areaUnderPR")
-    tvs = TrainValidationSplit(estimator=pipeline,
-                               estimatorParamMaps=paramGrid,
-                               evaluator=pr_evaluator,
-                               trainRatio=0.7,
-                               collectSubModels=True)
+    tvs = CrossValidator(estimator=pipeline,
+                         estimatorParamMaps=paramGrid,
+                         evaluator=pr_evaluator,
+                         collectSubModels=True)
 
     model = tvs.fit(train_samples)
 

### 2019-04-15 preprocess.py
Fix loc_lat and loc_long type

@@ -31,14 +31,15 @@ def preprocess_accidents(accidents_df):
             .withColumn('date', to_date(col('DT_ACCDN'), format='yyyy/MM/dd'))
             .withColumn("hour", split(col('HEURE_ACCDN'), ':')[0].cast("int"))
             .drop('DT_ACCDN', 'HEURE_ACCDN')
-            .withColumnRenamed('LOC_LAT', 'loc_lat')
-            .withColumnRenamed('LOC_LONG', 'loc_long')
+            .withColumn('loc_lat', col('LOC_LAT').astype('double'))
+            .withColumn('loc_long', col('LOC_LONG').astype('double'))
+            .drop('LOC_LAT', 'LOC_LONG')
             .withColumnRenamed('ACCIDENT_ID', 'accident_id')
             .dropna())
 
 
 def match_accidents_with_roads(spark, road_df, accident_df, use_cache=True):
-    cache_path = workdir + 'data/match_accident-road.parquet'
+    cache_path = workdir + 'data/matches_accident-road.parquet'
     if isdir(cache_path) and use_cache:
         print('Reading accident-road matches from cache...')
         return spark.read.parquet(cache_path)

### 2019-04-15 preprocess.py
Fix loc_lat and loc_long type

@@ -31,14 +31,14 @@ def preprocess_accidents(accidents_df):
             .withColumn('date', to_date(col('DT_ACCDN'), format='yyyy/MM/dd'))
             .withColumn("hour", split(col('HEURE_ACCDN'), ':')[0].cast("int"))
             .drop('DT_ACCDN', 'HEURE_ACCDN')
-            .withColumnRenamed('LOC_LAT', 'loc_lat')
-            .withColumnRenamed('LOC_LONG', 'loc_long')
+            .withColumn('loc_lat', col('LOC_LAT').astype('double'))
+            .withColumn('loc_long', col('LOC_LONG').astype('double'))
             .withColumnRenamed('ACCIDENT_ID', 'accident_id')
             .dropna())
 
 
 def match_accidents_with_roads(spark, road_df, accident_df, use_cache=True):
-    cache_path = workdir + 'data/match_accident-road.parquet'
+    cache_path = workdir + 'data/matches_accident-road.parquet'
     if isdir(cache_path) and use_cache:
         print('Reading accident-road matches from cache...')
         return spark.read.parquet(cache_path)

### 2019-04-15 preprocess.py
Fix RandomUnderSampler: now only undersampling training set

@@ -444,4 +444,4 @@ def get_dataset_df(spark, pos_samples, neg_samples):
                   'street_id',
                   'date', 'hour', 'features', 'label'))
 
-    return df
+    return df.withColumn('id', monotonically_increasing_id())

### 2019-04-15 random_undersampler.py

@@ -1,6 +1,7 @@
 from pyspark.ml.param.shared import Param, Params, HasSeed, HasLabelCol
-from pyspark.ml import Transformer
-from pyspark.sql.functions import col
+from pyspark.ml import Transformer, Estimator
+from pyspark.sql.functions import col, lit, isnull
+from pyspark.ml.param import TypeConverters
 
 
 class HasTargetImbalanceRatio(Params):
@@ -26,13 +27,52 @@ class HasTargetImbalanceRatio(Params):
         """
         return self._set(targetImbalanceRatio=value)
 
+class HasIndexCol(Params):
+    """
+    Mixin for param indexCol: index column name.
+    """
 
-class RandomUnderSampler(Transformer, HasTargetImbalanceRatio, HasSeed,
-                         HasLabelCol):
-    def _transform(self, dataset):
+    indexCol = Param(Params._dummy(), "indexCol", "index column name.", typeConverter=TypeConverters.toString)
+
+    def __init__(self):
+        super(HasIndexCol, self).__init__()
+        self._setDefault(indexCol='index')
+
+    def setIndexCol(self, value):
+        """
+        Sets the value of :py:attr:`indexCol`.
+        """
+        return self._set(indexCol=value)
+
+    def getIndexCol(self):
+        """
+        Gets the value of indexCol or its default value.
+        """
+        return self.getOrDefault(self.indexCol)
+
+
+class RandomUnderSampler(Estimator, HasTargetImbalanceRatio, HasSeed,
+                         HasLabelCol, HasIndexCol):
+    def _fit(self, dataset):
         neg_samples = dataset.filter(col(self.getLabelCol()) == 0.0)
         pos_samples = dataset.filter(col(self.getLabelCol()) == 1.0)
         current_ratio = neg_samples.count()/pos_samples.count()
         sampling = self.getTargetImbalanceRatio()/current_ratio
-        neg_samples = neg_samples.sample(sampling, seed=self.getSeed())
-        return neg_samples.union(pos_samples)
+        indexes_to_remove = neg_samples.select(self.getIndexCol()).sample(1-sampling)
+
+        return RandomUnderSamplerModel(indexes_to_remove, self.getSeed())
+
+class RandomUnderSamplerModel(Transformer):
+    def __init__(self, indexesToRemove, seed):
+        super(RandomUnderSamplerModel, self).__init__()
+        self.seed = seed
+        self.indexesToRemove = indexesToRemove
+
+    def _transform(self, dataset):
[…]

### 2019-04-15 main_gen_match.py
Fix main_gen_match.py

@@ -46,6 +46,12 @@ def generate_match_accident_road_of_one_month(year, month):
     match_accident_road.write.parquet(filepath)
     spark.stop()  # Force garbage collection and empty temp dir
 
+
+# Make sure to generate accidents DF
+spark = init_spark()
+get_accident_df(spark)
+spark.stop()
+
 for year in range(2012, 2018):
     for month in range(1, 13):
         generate_match_accident_road_of_one_month(year, month)

### 2019-04-15 main_random_forest.py
Change param grid for random forest tuning

@@ -13,18 +13,17 @@ from workdir import workdir
 
 spark = init_spark()
 
-neg_samples = get_negative_samples(spark).sample(0.01)
-pos_samples = get_positive_samples(spark).sample(0.01)
+neg_samples = get_negative_samples(spark).sample(0.05)
+pos_samples = get_positive_samples(spark).sample(0.05)
 df = get_dataset_df(spark, pos_samples, neg_samples)
 df_sample = df.filter(col('date') > datetime.fromisoformat('2017-01-01'))
 (train_set, test_set) = df_sample.randomSplit([0.7, 0.3])
 
 model = random_forest_tuning(train_set)
 
-with open(workdir + 'data/random_forest_tuning_results5.txt', 'w') as file:
-    for models, result in zip(model.subModels, model.avgMetrics):
+with open(workdir + 'data/random_forest_tuning_results_7.txt', 'w') as file:
+    for model, result in zip(model.subModels, model.validationMetrics):
         file.write('==================================\n')
-        model = models[0]
         for stage in model.stages:
             params = stage.extractParamMap()
             for k in params:
@@ -33,8 +32,8 @@ with open(workdir + 'data/random_forest_tuning_results5.txt', 'w') as file:
 
 predictions = model.transform(test_set)
 area_under_PR, f1_score = evaluate_binary_classifier(predictions)
-with open(workdir + 'data/random_forest_tuning_perf_5.txt', 'w') as file:
+with open(workdir + 'data/random_forest_tuning_perf_7.txt', 'w') as file:
     file.write(f"Area Under PR = {area_under_PR}\nF1 score = {f1_score}")
 
 graph = compute_precision_recall_graph(predictions, 20)
-graph.to_csv(workdir + 'data/precision_recall_graph_rf_5.csv')
+graph.to_csv(workdir + 'data/precision_recall_graph_rf_7.csv')

### 2019-04-15 random_forest.py

@@ -1,7 +1,7 @@
 from pyspark.ml.classification import RandomForestClassifier
 from pyspark.ml.evaluation import BinaryClassificationEvaluator
 from pyspark.ml.evaluation import MulticlassClassificationEvaluator
-from pyspark.ml.tuning import CrossValidator, ParamGridBuilder
+from pyspark.ml.tuning import TrainValidationSplit, ParamGridBuilder
 from pyspark.ml import Pipeline
 from pyspark.sql.types import DoubleType
 from pyspark.sql.functions import udf, col
@@ -15,25 +15,26 @@ def random_forest_tuning(train_samples):
     rf = RandomForestClassifier(labelCol="label",
                                 featuresCol="features",
                                 cacheNodeIds=True)
-    ru = RandomUnderSampler()
+    ru = RandomUnderSampler().setIndexCol('id')
     pipeline = Pipeline().setStages([ru, rf])
     paramGrid = \
         (ParamGridBuilder()
-         .addGrid(rf.numTrees, [100, 115, 130])
+         .addGrid(rf.numTrees, [50, 75, 100])
          .addGrid(rf.featureSubsetStrategy, ['sqrt'])
-         .addGrid(rf.impurity, ['entropy'])
+         .addGrid(rf.impurity, ['gini', 'entropy'])
          .addGrid(rf.maxDepth, [30])
          .addGrid(rf.minInstancesPerNode, [1])
-         .addGrid(rf.subsamplingRate, [0.8, 0.6])
-         .addGrid(ru.targetImbalanceRatio, [1.0, 1.5, 1.25])
+         .addGrid(rf.subsamplingRate, [0.6, 0.5, 0.4])
+         .addGrid(ru.targetImbalanceRatio, [1.0, 1.5, 2.0])
          .build())
     pr_evaluator = \
         BinaryClassificationEvaluator(labelCol="label",
                                       rawPredictionCol="rawPrediction",
                                       metricName="areaUnderPR")
-    tvs = CrossValidator(estimator=pipeline,
+    tvs = TrainValidationSplit(estimator=pipeline,
                          estimatorParamMaps=paramGrid,
                          evaluator=pr_evaluator,
+                         trainRatio=0.7,
                          collectSubModels=True)
 
     model = tvs.fit(train_samples)

### 2019-04-15 random_undersampler.py

@@ -58,11 +58,13 @@ class RandomUnderSampler(Estimator, HasTargetImbalanceRatio, HasSeed,
         pos_samples = dataset.filter(col(self.getLabelCol()) == 1.0)
         current_ratio = neg_samples.count()/pos_samples.count()
         sampling = self.getTargetImbalanceRatio()/current_ratio
-        indexes_to_remove = neg_samples.select(self.getIndexCol()).sample(1-sampling)
+        indexes_to_remove = (neg_samples.select(self.getIndexCol())
+                             .sample(1-sampling, seed=self.getSeed())
 
-        return RandomUnderSamplerModel(indexes_to_remove, self.getSeed())
+        return (RandomUnderSamplerModel(indexes_to_remove, self.getSeed())
+                .setTargetImbalanceRatio(self.getTargetImbalanceRatio()))
 
-class RandomUnderSamplerModel(Transformer):
+class RandomUnderSamplerModel(Transformer, HasTargetImbalanceRatio):
     def __init__(self, indexesToRemove, seed):
         super(RandomUnderSamplerModel, self).__init__()
         self.seed = seed

### 2019-04-15 presentation/pres.tex
adding references, fixing typos, finalizing slides

@@
   \begin{itemize}
-    \item Extensive use of decision trees in several form. (Athanasios Theofilatos, A. 2017, Joaquín Abellán, 2013, Lei Lin 2014, Li-Yen Chang 2005)
-    \item Recently, deep learning algorithms (LSTM and CNN architectures for example) (Zhuoning Yuan, 2018, Quanjun Chen 2016)
+    \item Extensive use of decision trees in several form. (\cite{Theofilatos2017}, \cite{Abellan2013}, \cite{Lin2015}, \cite{Chang2005})
+    \item Recently, deep learning algorithms (LSTM and CNN architectures for example) (\cite{Yuan2018}, \cite{Chen2016})
   \end{itemize}
@@
     \item Every time an accident did not happen is considered a negative sample.
-		\item A sample from the cartesian product between all dates (between 01/01/2012 and 31/12/2017) and all roads.
+		\item We extracted a subsample from the cartesian product between all dates (between 01/01/2012 and 31/12/2017) and all roads.
 		\item Generation of 5 million negative samples for about 200 000 positive samples.
@@
 	\begin{itemize}
-		\item Computing the distances between GPS coordinates (taking into account earth curvature)
-		\item Making the dates and times cyclic
-		\item Creating missing points on the roads
-    \item Interpolating weather statistics by averaging data from several stations
+		\item Computation of the distances between roads and accidents using their GPS coordinates (taking into account earth curvature)
+		\item Transformation of the dates and times into a cyclic format
+		\item Extrapolation of missing points on the roads (defined by a limited number of points in a KML format)
+    \item Interpolation of weather statistics at a given point by averaging data collected from several stations
     \item etc...
@@
 		\end{equation*}
-    \item Data imbalance imply decrease of TP and increase of FN
-    \item Therefore, area under PR exhibits larger difference than are under ROC
+    \item Data imbalance imply decrease of TP and increase of FN (none of them appears in the FPR)
+    \item Therefore, area under PR exhibits larger differences than area under ROC (\cite{Davis2006})
 	\end{itemize}
@@
     	\item Each tree is trained on a random sample of the data and a random sample of the features.
+      \item The final prediction is given by averaging the votes of the trees.
     \end{itemize}
@@
     \begin{itemize}
-      \item A solution to imbalance.
+      \item A solution to imbalance (\cite{Chen2004}).
       \item Add weights to the classes when growing the tree (in the Gini coefficient).
@@
     \begin{itemize}
-      \item Another solution to imbalance.
-      \item Combine undersampling and ensemble method versus loss of information.
-      \item For each tree : Randomly draw sample of same size from majority (with replacement) and minority class.
+      \item Another solution to imbalance (\cite{Chen2004}).
+      \item Combine undersampling and ensemble method to migitate the loss of information.
+      \item For each tree : Randomly draw samples of same size from majority (with replacement) and minority class.
       \item The rest of the algorithm is identical to Random Forests
@@
       \item No clear winner ! But...
-      \item WRF - More vulnerable to noise on minority class
-      \item BRF - More computationally efficient because use less data (subsampling)
-      \item BRF - Easier to implement
+      \item WRF - More vulnerable to noise on minority class (\cite{Chen2004})
+      \item BRF - More computationally efficient because use less data (subsampling) (\cite{Chen2004})
+      \item BRF - Easier to implement in Spark (see next slide)
     \end{itemize}
@@
     	\item To know how many times a given sample will be in the subsample of a given tree (can be 0 times), we use "distribution sampling" of 1 element.
-      \item (Distribution sampling: Get a random sample following a given distribution)
+(Distribution sampling: Get a random sample following a given distribution)
     \end{itemize}
@@
   \item Area Under PR : 99,3\% for WRF VS 99,7\% for BRF
-  \item F1 score : 99\% for WRF VS 98\% for BRF (which proves that F1 score can be misleading)
+  \item F1 score : 99\% for WRF VS 98\% for BRF (which proves that F1 score can be misleading in the context of class imbalance)
 
@@
         \item Spark's map-reduce and shuffling operations are very handy
-        \item Spark ML pipelines allow to create pipelines of preprocessing easily
+        \item Spark ML pipelines allow to create processing pipelines easily
       \end{itemize}
@@
 		\item The most important features seem to be the time and location of the vehicle, followed by some road's features like the type of road (highway for example) and the street length.
-    \item Finally, comes wome weather features like the visibility, the wind and the humidity.
+    \item Finally, come some weather features like the visibility, the wind and the humidity.
 	\end{itemize}
@@
     \item We identified some issues related to class imbalance and big data analytics.
-    \item Future work would include tuning the XGBoost classifier, testing lightGBM classifier, assing some datasets like a road works dataset.
+    \item We implemented BRF in Spark.
+    \item Future work would include tuning the XGBoost classifier, testing lightGBM classifier, adding some datasets like a road works dataset.
     \item The code can be found on Github at https://github.com/GTimothee/accident-prediction-montreal

### 2019-04-15 presentation/pres.tex
add figures and results

@@
   \begin{itemize}
-    \item Road accidents implies millions of people deaths and injuries every year.
+    \item Road accidents result in millions of people deaths and injuries every year.
 		\item Human and economic impact on society.
@@
   \begin{itemize}
-    \item Dataset analysis.
+    \item Dataset assembling.
+    \item Predict the risk of road accident at a given time on a given road.
   	\item Finding key insights in accident prediction.
@@
     \item Shape of the road, how roads are
-connected together, the mail addresses on the street, the number of lanes in each direction and the
-type of pavement.
+connected together, the mail addresses on the street, etc.
   \end{itemize}
@@
   \begin{itemize}
-    \item For each position and hour, the dataset provides the temperature, the humidity, the wind speed and direction, the visibility, the atmospheric pressure etc.
+    \item The dataset provides the temperature, the humidity, the wind speed and direction, the visibility, the atmospheric pressure etc. at different weather stations in Montreal.
   \end{itemize}
@@
 \begin{frame}
-	\frametitle{Generating the negative samples}
+	\frametitle{Generating the examples}
+  \begin{itemize}
+  \item[] \textbf{Positive samples}:
+    \begin{itemize}
+      \item Identification of road for each accident: Computation of the distances between roads and accidents using their GPS coordinates (taking into account earth curvature)
+		  \item Extrapolation of missing points on the roads for more precise road identification
+    \end{itemize}
+  \item[] \textbf{Negative samples}:
     \begin{itemize}
@@
   		\item Generation of 5 million negative samples for about 200 000 positive samples.
+    \end{itemize}
+
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Computation of the distances between roads and accidents using their GPS coordinates (taking into account earth curvature)
 		\item Transformation of the dates and times into a cyclic format
-		\item Extrapolation of missing points on the roads (defined by a limited number of points in a KML format)
     \item Interpolation of weather statistics at a given point by averaging data collected from several stations
-    \item etc...
+    \item Road features: street type and length
 	\end{itemize}
@@
     \begin{itemize}
-      \item No clear winner ! But...
+      \item No clear winner in terms of prediction power ! But...
       \item WRF - More vulnerable to noise on minority class (\cite{Chen2004})
@@
 \begin{itemize}
-  \item Area Under PR : 99,3\% for WRF VS 99,7\% for BRF
-  \item F1 score : 99\% for WRF VS 98\% for BRF (which proves that F1 score can be misleading in the context of class imbalance)
+  \item Area Under PR : 98,1\% for RF VS 98,7\% for BRF
+  \item F1 score : 99.7\% for RF VS 98.8\% for BRF (which proves that F1 score can be misleading in the context of class imbalance)
+  \item With BRF, we detect more accidents but it increases the false positive rate.
 
@@
 \centering
-\includegraphics[height=3.5cm, keepaspectratio]{Figures/BRF.png}
-\includegraphics[height=3.5cm, keepaspectratio]{Figures/WRF.png}
+\includegraphics[height=3.0cm, keepaspectratio]{Figures/RF.png}
+\includegraphics[height=3.0cm, keepaspectratio]{Figures/BRF.png}
 
@@
 	\begin{itemize}
-		\item Using slurm-based (scheduler) cluster to increase the dataset generation's speed.
+		\item Using cluster to increase the dataset generation's speed.
     \item Tricks to increase the generation speed and reduce the memory consumption.
@@
 \centering
-\includegraphics[height=3.5cm, keepaspectratio]{Figures/timeVSsamples_generatepositivesamples.png}
-\includegraphics[height=3.5cm, keepaspectratio]{Figures/timeVSsamples_generatepositivesamples_after.png}
+\includegraphics[height=5.0cm, keepaspectratio]{Figures/timeVSsamples_generatepositivesamples.png}
 \end{frame}
@@
         \item Spark's map-reduce and shuffling operations are very handy
-        \item Spark ML pipelines allow to create processing pipelines easily
       \end{itemize}
@@
 \begin{frame}
-	\frametitle{Comparison between the algorithms}
+	\frametitle{Results}
 	\begin{itemize}
-		\item Random forest
-		\item Balanced random forest
-		\item XGBoost
+		\item Random forest: 15.3\% area under PR curve
+		\item Balanced random forest: 17.2\% area under PR curve
 	\end{itemize}
+\centering
+\includegraphics[height=4cm, keepaspectratio]{Figures/rf_results.png}
+\includegraphics[height=4cm, keepaspectratio]{Figures/nohyperparam_brf.png}
 \end{frame}
@@
     \item We found some key features that can explain the occurence of accidents.
-    \item We identified some issues related to class imbalance and big data analytics.
+    \item We identified some solution to deal with class imbalance and big data-related issues.
     \item We implemented BRF in Spark.
-    \item Future work would include tuning the XGBoost classifier, testing lightGBM classifier, adding some datasets like a road works dataset.
+    \item Future work would include tuning the XGBoost classifier, adding some datasets like a road works dataset.
     \item The code can be found on Github at https://github.com/GTimothee/accident-prediction-montreal

### 2019-04-16 preprocess.py
Fix one hot encoding of dayofweek

@@ -375,7 +375,7 @@ def add_date_features(samples):
     samples = add_cyclic_feature(samples, month('date'), 'month', 12)
     samples = add_cyclic_feature(samples, dayofmonth('date'), 'day', 31)
 
-    samples = samples.withColumn('dayofweek', dayofweek('date'))
+    samples = samples.withColumn('dayofweek', dayofweek('date') - 1)
     encoder = OneHotEncoder(inputCols=['dayofweek'],
                             outputCols=["dayofweek_onehot"])
     encoder_model = encoder.fit(samples)

### 2019-04-18 paper/paper.tex
Add paper draft

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Road Accident Prediction in Time and Space}

\author{\IEEEauthorblockN{1\textsuperscript{st} Antoine Hébert}
\IEEEauthorblockA{\textit{Department of Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montréal, Québec, Canada \\
an\_heb@encs.concordia.ca}
\and
\IEEEauthorblockN{1\textsuperscript{st} Timothée Guédon}
\IEEEauthorblockA{\textit{Department of Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montréal, Québec, Canada \\
t\_guedon@encs.concordia.ca}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Tristan Glatard}
\IEEEauthorblockA{\textit{Department of Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montréal, Québec, Canada \\
tglatard@encs.concordia.ca}
\and
\IEEEauthorblockN{4\textsuperscript{th} Brigitte Jaumard}
\IEEEauthorblockA{\textit{Department of Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montréal, Québec, Canada \\
bjaumard@cse.concordia.ca}}

\maketitle

\begin{abstract}
Road accidents are an important issue of our modern societies, responsible for millions of deaths and injuries. In this paper we show how one can leverage open datasets from a City like Montreal, Canada, to create some accident prediction models, using state-of-the-art big data analytics methods in Python. Such models could then be used in the context of road accident prevention, but also to identify key factors that can lead to a road accident, and eventually, help to elaborate new policies. This study also explains how we dealt with the severe class imbalance issue of the accident prediction problem, a recurrent issue that touches many other scientific domains like medical diagnoses or fraud detection. In particular, we show how we implemented Balanced Random Forests, a variant of the Random Forests machine learning algorithm, into the Scala and Python APIs of Apache Spark big data framework. All source code can be found on Github.
\end{abstract}

\begin{IEEEkeywords}
machine learning, road accidents prediction
\end{IEEEkeywords}

\section{Introduction}
\subsection{Context}
According to the World report on road traffic injury prevention, 2004, more than one million people die every year from car accidents and dozens of millions of people are being injured\cite{Peden2004}. In this report, the World Health Organization also describe road traffic systems as “the most complex and the most dangerous system with which people have to deal every day”. This issue is all the more important that the number of accidents is forecast to continue rising. In the last decade, Big Data Analytics has been emerging \cite{Gandomi2015}, allowing scientists to handle a very large amount of data, but also complex and heterogeneous data, in order to extract useful insights from it. In the context of accident prediction, Big Data Analytics could provide insights on the conditions leading to an increased risk of road accidents that could then be used to develop traffic-related policies and prevention operations. Our contributions include: (1) a demonstration of how simple and easily available datasets can be assembled to obtain meaningful features for road accident prediction, (2) a comparison of three different machine learning algorithm dealing with data imbalance in the context of road accident prediction, (3) the implementation of Balanced Random Forest introduced by Chen et al. in Apache Spark for efficient distributed training. Finally, all the source code used is publicly available on Github under the MIT license.

\subsection{Related work}
Accident prediction has been extensively studied in the last decade. The main techniques that have been used at the beginning of the research in road accidents prediction were methods using the Poisson distribution and the negative binomial regression \cite{Chang2005}. Then, as machine learning methods become more popular, algorithms like decision trees have been used in several forms like classification and regression decision trees, frequent pattern trees and random forests
\cite{Chang2005, Theofilatos2017, Abellan2013, Lin2015}, along with other machine learning algorithms like Support Vector Machines \cite{LI20081611}. Recently, the outstanding progress of deep learning algorithms have resulted in their extensive use in every field, including in Geospatial machine learning (Convolutional Neural Network and Long Short-Term Memory neural network for example) \cite{Yuan2018, QChen2016}.
\section{Methods}
\subsection{Choosing a Big Data framework: Apache Spark or Dask}
According to Dask’s official website, “Dask is a flexible library for parallel computing in Python” \cite{dask}. Apache Spark, on the other hand, describes itself as “a fast and general-purpose cluster computing system” \cite{spark}. Both of them are accepted tools with wide supporting community, even though Apache Spark is a bigger and older project. Although we almost exclusively used Apache Spark’s Python API (called “pySpark”) in this study, we started our experiment using Dask. Some advantages of Dask over Spark made us chose it for our analysis in the first place; For example, Dask “is developed in coordination with […] projects like Numpy, Pandas and Scikit-Learn” \cite{dask}, all widely used Python packages in machine-learning and data science. It is also admitted that Dask is lighter than Spark and designed to be more flexible in terms of applications and algorithms. However, we found ourselves writing a lot more code using Dask than Spark for the same result. Spark map-reduce-based interface is very handy and allow easier processing of datasets. For all these reasons we found out that Spark fitted our needs best and we switched to it at an early time of the study.

\subsection{Algorithms selected}
For this study, we chose to focus on tree-based algorithms because they have proven their efficiency compared to classical statistical methods and allow for easier interpretability than deep learning algorithms. In particular, we implemented the Balanced Random Forests (BRF) algorithm in Apache Spark to deal with class imbalance which is a prominent issue in road accident prediction. To validate the performance of our version, we run an experiment with a simple synthetic dataset to compare its performances with the standard implementation of Random Forest (RF) in Spark with random under-sampling of the majority class. Balanced Random Forest performed much better. Balanced Random Forests in one of the two methods proposed by Chen, Liaw, and Breiman to deal with class imbalance when using Random Forest. Balanced Random Forests are like Random Forests, but with a difference during the bootstrapping phase, for each tree of the forest, a random under-sampling of the majority class is performed in order to obtain a balanced sample. Intuitively, Balance Random Forest is an adaptation of random undersampling of the majority class making use of the fact that Random Forests are an ensemble method. Random undersampling usually performs better than more advanced methods like SMOTE or NearMiss \cite{Branco2016}. Chen, Liaw and Breiman\cite{Chen2004} also proposed in the same paper, the Weighted Random Forest (WRF) which consists in giving more weight to the minority class at the building time of each tree and during the measure of the gain in impurity and during the class prediction of each terminal node. While neither method is clearly better than the other in terms of prediction power, we chose BRF because it is found to be more computationally efficient (because of the under-sampling) and easier to implement in our opinion. Interestingly, Wallace et al. \cite{Wallace2011} presents a theoretical analysis of the data imbalance problem and suggest to use methods like Balance Random Forest.

We decided to compare the performance of the Balanced Random Forests algorithm with the performances of one of the very efficient implementations of Gradient Boosted Trees: XGBoost which offers parameters to deal with data imbalance as well.

\subsection{Datasets}
We used three public datasets provided by the city of Montreal and the government of Canada: (1) Montreal Vehicle Collision: This dataset provided by the city of Montreal contains all the road collisions reported by the police in Montreal from 2012 to 2018. For each accident, the dataset contains the date and localization of the accident, information on the number of injuries and death, the number of vehicles involved, and information on the road condition. We used only the localization and the date of the accident since we do not have the other piece of information when no accident happened. Another dataset with all vehicle collision in Canada is available but without the localization of the accident, therefore we restrained our analysis to the city of Montreal. (2) National Road Network: This dataset provided by the government of Canada contains the geometry of all roads in Canada. For each road segment, a few meta-data are given. Sadly, for roads in Quebec the speed limit and the surface type are not provided. The data was available in various format, we chose to use the Keyhole Markup Language, which is easy to read because it is based on the Extensible Markup Language (XML). (3) Data Historical Climate Data: This dataset provided by the government of Canada contains hourly weather information measured at different weather stations across Canada. For each station and every hour, the dataset provides the temperature, the humidity, the wind direction and speed, the visibility, the atmospheric pressure and observations of atmospheric phenomenon such as snow, fog, rain, etc.

\subsection{Positive and negative samples generation}
Our machine learning problem can be stated as a binary classification problem, the positive class being the occurrence of an accident and the negative class being all the cases where an accident did not happen. We got a little bit less than 150 000 positive samples from the national collision dataset and we generated a random sub-sample of 5 million negative samples (over 10 billion possibilities). The generation of the negative samples basically consisted in a Cartesian product between the roads from the road network dataset and all the hours of all days that occurred between the 1st January 2012 and the 31 of December 2017, the dates range of the Montreal Vehicle Collision dataset. This huge cartesian product operation together with the joins with the data from the other datasets made our dataset generation pretty heavy. It resulted in an impractical amount of time and memory space requirements to generate the dataset. Our first implementation was querying the Historical Climate Data API in real-time with a cache mechanism. Our idea was to collect only the weather stations and hours necessary for our sample of negative examples, but it resulted in bad performances. We got a performance increase by first building a Spark data frame with all the Historical Climate Data for weather stations around Montreal and then joining the two datasets. We conducted a detailed analysis of our algorithm to improve its performances. We notably obtained a good performance increase by not persisting intermediary results of the road segment identification for accidents. As opposed to what we initially thought, recomputing these results was faster than writing and reading them in the cache. Finally, the identification of the road segment corresponding to accidents was very memory intensive, we modified this step to be executed by batch of one month.  With these improvements and a few other tricks including partitioning the data frame at key points in our algorithm, we manage to reduce the processing time from about one minute to generate about 1000 positive samples to less than 1 second. We also used a cluster from “Compute Canada” to take maximum advantage of Apache Spark distributed nature for the generation of examples and the hyper-parameter tuning of our models.

\subsection{Feature Engineering}
Once we had our pairs of date with hour and road segment for positive and negative examples, we created features from this information. We created four types of features, weather features, features from the road segment, features from the date .

For the weather features, we used data from the Historical Climate Dataset. In order to estimate the weather information at the position of the road segment we compute the mean of the weather information from all  the surrounding weather stations at the date and hour of the example  weighted by the inverse of the square of the distance between the station and the road segment. We initially used the inverse of the distance, but we obtained a small improvement in performances when squaring the inverse of the distance. It makes sense since, the weather information is probably much closer to the closest stations than the other ones. We tried higher power, but results were not as good. It could be interesting to experiment with different power values for each weather information. We used all the weather information provided by the Historical Climate Dataset. The features are: the temperature, the dew point temperature which is a measure of the humidity, the real humidty percentage, the wind direction, the wind speed, the visibility, the atmospheric pressure, the Hmdx index, which is a measure of felt temperature and the wind chill which is another measure of the felt temperature using the wind information. In addition, the historical weather dataset, provide a multi-label categorical variable providing the observations of atmospheric phenomenon such as snow, fog, rain, etc. We used this last variable to create a binary variable that we called risky weather which is true when the following phenomenon are observed: Freezing Rain, Freezing Drizzle, Snow, Snow Grains, Ice Crystals, Ice Pellets, Ice Pellet Showers, Snow Showers, Snow Pellets, Ice Fog, Blowing Snow, Freezing Fog.

For the features from the road segments, we were limited by the limited metadata provided on the road segments. From the shape of the road segment, we computed the length of the road segment, and from the name of the street we identified the type of road (Highway, street, boulevard, etc.). In addition road segment are classified in three different level in the dataset depending on their importance in the road network, we created a categorical feature from this information. For these two categorical features, we encoded them as suggested in Element of Statistical Learning\cite{elementsofstat} in section 9.2.4, instead of using one hot encoding which would create an exponential number of splits, we index the categorical variable ordered by the proportion of the example belonging to the given category which are positive samples. This encoding guarantee to provide optimal splits on these categorical variables. Lastly, we added an a feature giving the number of accidents which occurred previously on this road segment.

For the date features, is is quite simple, we took the day of the year, the hour of the day, and the day of the week. We decided to encode day of the year and hour of the day to be cylic. Indeed, for example with hour of the day, 23 hour is very close to 1 hour, but in the usual encoding this does not appear to the machine learning algorithm. With cyclical encoding, we compute two features, the first one is the cosine of the original feature scaled between 0 and two pi, and the second one is the sine of the original feature scaled between 0 and two pi.

\subsection{Implementation of balanced random forest in Apache Spark}
The Balanced Random Forests algorithm was not implemented in Apache Spark. An implementation is available in the python library imbalanced-learn\cite{imbalance} which implements many algorithm to deal with data imbalance using an API inspired by scikit-learn, but the size of our dataset made it impossible for us to use this library. Therefore, we implemented Balance Random Forests in Apache Spark.

In the Apache Spark implementation of Random Forests, the bootstrap step is made before starting to grow any tree. For each sample, an array contains the number of time it will appear in each tree. When doing sampling with replacement, each value of this array is given by a Poisson distribution. The parameter of the Poisson distribution correspond to the subsampling rate hyper-parameter of the Random Forest, which specifies the size of the sample used for training each tree as a fraction of the total size of the dataset. Indeed, if for example we want each tree to use a sample of the same size of the whole dataset, the subsampling ratio will be set to 1.0, which is indeed the number of time a given example will appear in a tree on average.

In order to implement Balanced Random Forests, we modified the parameter of the Poisson distribution to use the class weight multiplied by the subsampling ratio, this way a negative sample with for example the weight 0.25 has 4 times less chance to be chosen to appear in a given tree. This implementation has the advantage that it did not required a big code change and will be easy to test, but it also has a disadvantage, users probably expect linearly correlated weight to be equivalent, which is not the case in our implementation.

In order to be compatible with other possible use cases, the weight are actually applied per samples and not per class. This is a choice made by Apache Spark developers that we respected. In order to support sample weights, we create a new Poisson distribution for each sample. To make sure the random number generator is not reseeded for each sample, we use the same underlying random number generator for all Poisson distributions, this also helps reduce the cost of creating a new Poisson distribution object.
Like with other estimators accepting weights, our Balanced Random Forests implementation use weight from a weight column in the samples data frame. We adapted the Python wrapper of the random forest classifier to accept and forward weights to the algorithm in Scala. Finally, we did a first validation of our implementation using an artificially generated imbalanced dataset and compared the results of both standard random forest and BRF. While BRF had better results in terms of area under PR (which was our selected measure) (98.\% against 98.1\%), standard random forest had better F1 score (99.7\% against 98.8\%) with default threshold. This emphasized that we were right in our choice of the area under PR curve as our measure in the case of class imbalance, indeed with a good threshold, the BRF model performs better.

\subsection{Hyper-parameter tunings}
In order to determine the optimal hyper parameter, we first performed automatic hyper parameter tuning using Apache Spark ML library grid search using cross validation. Because the training on the whole dataset would have been too high, we took a small sample of the dataset. Still, We could not test many parameters combination using this method.

Once we got a first tuning result with grid search we continued manually by following a plan, do , check, adjust method, using the measure of the area under the precision recall curve, the precision and recall with different threshold on the test set and the training set we were able to better understand how the performances of our model could be improved.

Interestingly, despite using many trees, our Random Forest algorithms tended to overfit very quickly as soon as the maximum depth parameter goes above 18. We eventually used only 100 trees, because more trees did not help much. We have not tried more than 200 trees, maybe many more trees would have been necessary to increase the maximum depth without over fitting, but then the training time would not be reasonable.

\subsection{Choice of a metric to evaluate our models}
The Chosen metric for the evaluation of the models was the area under Precision/Recall (PR) curve (area under PR). According to Davis and Goadrich\cite{Davis} this measure is the most representative of how well a model has been trained in the case of data imbalance, particularly compared to the ROC curve which is the commonly used measure. One can intuitively be convinced because in the ROC curve the false positive rate (FPR) is used instead of the precision. The precision refers to the proportion of examples detected as positive which are actually positive, while the FPR refers to the proportion of negative detected as positives. With an imbalanced dataset, the FPR will usually stay very low because the number of negative is high. The precision, by contrast, will usually be quite low since algorithms tuned for data imbalance will usually include many false positives in the examples they detect in order to achieve a correct recall also called true positive rate.

\subsection{Identifying the most important features}
Random forests allow to measure feature importances by computing the total decrease in impurity of all splits using the variable weighted by the number of samples. This variable importance measure is not perfect for interpretability since it is biased toward non-correlated variables, but it helps to select the most useful features for the prediction. You can find the feature importances given by the Balance Random Forests in figure 1. According to this figure, the most important features to predict the occurrence of an accident is the number of accidents which occurred on this road during the previous years which we could have easily guessed. Then the location of the road, the type and length of the road segment and the hour of the day have a high importance. As compared to the count of accident feature the remaining features seem to have almost no importance, but the model does not perform as well if we remove them. Surprisingly, the day of the year, the temperature and the risky weather boolean have very low feature importance. We believe that the weather features have lower feature importances because they are strongly correlated. The gain in impurity the information they contain provide is distributed over all the correlated features. Also, the risky weather feature would probably require more feature engineering, since if it was snowing one hour ago, there is still a higher risk. We could do a moving average.

\begin{figure}[htbp]
\centerline{\includegraphics[height=7cm, keepaspectratio]{figures/feature_importances.png}}
\caption{Feature importances computed by the Balanced Random Forest}
\label{fig}
\end{figure}

\section{Results}
The training time of the machine learning algorithm is very long due to the size of our dataset and the latest improvement in term of feature engineering have been discovered very late. Therefore, we currently only have the results with the Balanced Random Forest using the latest features. The results from the previous training of other algorithm are not comparable since they do not use the same features. We will add the missing results in later versions of this document.

These results have been obtained by training the Balanced Random Forest algorithm on the whole dataset of positive samples, but with only a fifth of the dataset of negative samples we have generated. Therefore the data imabalance is reduced to a factor of 8. According to our previous experiments, the results are almost the same when using the full dataset of negative samples, sometimes slightly better. To evaluate this model we used a test set containing the last two years of our dataset, the model is trained on the 4 previous years and use only data from these years. For example, the "count\_accident" feature contains only the count of accidents on road segments which occurred from 2012 to 2016.

We obtain an area under the precision-recall curve of 79\%, and an area under the ROC curve of 97\%.
This results in the following precision and recall values depending on the chosen probability threshold:

\begin{figure}[htbp]
\centerline{\includegraphics[height=6cm, keepaspectratio]{figures/pr.png}}
\caption{Precision and recall as a function of the probablity threshold}
\label{fig}
\end{figure}

As we can see, with a threshold set to 0.63, we can achieve a recall of 85\% with a precision of 0.52\%. With a threshold of 0.06, the model can actually detect almost all the accidents, 94\% of them, with a precision of 37\%, this corrsponds to the high-risk roads and time in which we predict a risk of 37\% of accidents.

 Our source code is available on Github at \href{https://github.com/GTimothee/accident-prediction-montreal}{https://github.com/GTimothee/accident-prediction-montreal}.

\section{Discussion}
The overall results are quite good but mostly rely on the count of previous accidents on the road segment. This is not an issue for accident prediction, but it does not help to understand why these roads are particularly dangerous. Accident prediction is a very hard machine learning problem because even if the risk is very high for an accident to occur, the driver might manage to avoid it in time and avoid the accident although the conditions were almost exactly the same as when an accident occurs. We would need information on the state of the drivers and the state of the vehicle to significantly improve performances.

However, we believe these results can be improved even without this important feature by adding more spatial features. We can see that one of the most important variables are the localization of the accident, but the ability to learn a spatial distribution by the random forest algorithm is limited. Therefore we believe providing more context information on where the road is located will greatly help. On another experiment we have seen the population density and the distance to the closest billboard used as features successfully. However, it will be hard to reach good performances because of the nature of the problem without the count of accidents.

\subsection{Future work}
We believe better performances could be reached by adding more features from other datasets. For the city of Montreal, we identified two particularly interesting datasets: a dataset with the location and dates of construction work on roads, and a dataset with the population density. These datasets might not be as easily available for over areas, but for Montreal, it could improve performances. Also, the solar elevation and the solar azimuth could easily be added without the need for another dataset. These features have been used successfully in road accident prediction in previous experiments.

\section*{Acknowledgment}

We acknowledge Compute Canada for giving us access to the cluster.

\bibliographystyle{IEEEtran}
\bibliography{biblio}

\end{document}

### 2019-04-18 preprocess.py
Improve date features: no onehot for dayofweek, replace month by dayofyear, make hour cyclic

@@ -6,8 +6,8 @@ from functools import reduce
 from pyspark.sql import Window, DataFrame
 from pyspark.sql.functions import row_number, col, rank, avg, split, to_date, \
                                   monotonically_increasing_id, when, isnan, \
-                                  year, cos, sin, month, dayofmonth, lit, \
-                                  dayofweek, isnull
+                                  year, cos, sin, dayofyear, dayofmonth, \
+                                  lit, dayofweek, isnull
 from pyspark.sql.types import BooleanType
 from pyspark.ml.feature import OneHotEncoder
 from pyspark.ml.feature import StringIndexer
@@ -372,19 +372,19 @@ def add_cyclic_feature(df, column, col_name, period):
 
 
 def add_date_features(samples):
-    samples = add_cyclic_feature(samples, month('date'), 'month', 12)
-    samples = add_cyclic_feature(samples, dayofmonth('date'), 'day', 31)
+    samples = add_cyclic_feature(samples, dayofmonth('date'), 'dayofmonth', 31)
+    samples = add_cyclic_feature(samples, dayofyear('date'), 'dayofyear', 366)
 
-    samples = samples.withColumn('dayofweek', dayofweek('date') - 1)
-    encoder = OneHotEncoder(inputCols=['dayofweek'],
-                            outputCols=["dayofweek_onehot"])
-    encoder_model = encoder.fit(samples)
-    samples = encoder_model.transform(samples).drop('dayofweek')
+    samples = samples.withColumn('dayofweek', dayofweek('date'))
+    samples = samples.withColumn('hour2', col('hour'))
+    samples = add_cyclic_feature(samples, col('hour'), 'hour', 24)
+    samples = samples.withColumnRenamed('hour2', 'hour')
 
     return samples
 
 
-features_col = ['hour',
+features_col = ['hour_cos',
+                'hour_sin',
                 'loc_long',
                 'loc_lat',
                 'street_level_indexed',
@@ -399,11 +399,12 @@ features_col = ['hour',
                 'wind_chill',
                 'hmdx',
                 'temp',
-                'month_cos',
-                'month_sin',
-                'day_cos',
-                'day_sin',
-                'dayofweek_onehot']
+                'dayofyear_cos',
+                'dayofyear_sin',
+                # 'dayofweek',
+                # 'dayofmonth_cos',
+                # 'dayofmonth_sin',
+                ]
 
 
 def remove_positive_samples_from_negative_samples(neg_samples, pos_samples):

### 2019-04-18 random_forest.py
Improve get_feature_importances to give features ordered by decreasing importance

@@ -99,4 +99,5 @@ def get_feature_importances(model):
     feature_names = features_col[:-1] + dayofweek_features
     feature_importances.index = feature_names
     feature_importances.columns = ["Feature importances"]
+    feature_importances.sort_values(['Feature importances'], ascending=False)
     return feature_importances

### 2019-04-18 random_undersampler.py
Fix random_undersampler.py

@@ -59,7 +59,7 @@ class RandomUnderSampler(Estimator, HasTargetImbalanceRatio, HasSeed,
         current_ratio = neg_samples.count()/pos_samples.count()
         sampling = self.getTargetImbalanceRatio()/current_ratio
         indexes_to_remove = (neg_samples.select(self.getIndexCol())
-                             .sample(1-sampling, seed=self.getSeed())
+                             .sample(1-sampling, seed=self.getSeed()))
 
         return (RandomUnderSamplerModel(indexes_to_remove, self.getSeed())
                 .setTargetImbalanceRatio(self.getTargetImbalanceRatio()))

### 2019-04-18 preprocess.py
Add risky_weather feature based on multi-label atmospheric phenomenon column of historical weather dataset

@@ -399,6 +399,7 @@ features_col = ['hour_cos',
                 'wind_chill',
                 'hmdx',
                 'temp',
+                'risky_weather',
                 'dayofyear_cos',
                 'dayofyear_sin',
                 # 'dayofweek',

### 2019-04-18 weather.py

@@ -8,6 +8,8 @@ from pyspark.sql.types import StructField, FloatType, StructType, \
                               IntegerType, ArrayType, DateType
 from requests import get
 import pandas as pd
+import numpy as np
+import re
 from utils import get_with_retry as get
 from workdir import workdir
 
@@ -19,7 +21,8 @@ COLUMNS_USED = ['Dew Point Temp (°C)',
                 'Stn Press (kPa)',
                 'Hmdx',
                 'Wind Chill',
-                'Temp (°C)']
+                'Temp (°C)',
+                'Weather']
 
 
 def get_weather_df(spark, accident_df):
@@ -96,6 +99,15 @@ def get_station_weather_month(station_id, year, month):
            f'&Month={month}&Day=14&'
            f'timeframe=1&submit=Download+Data')
     df = get_pandas_dataframe(url)[['Day', 'Time']+COLUMNS_USED]
+
+    # When temperature is not available, it means no weather data is
+    # available at all, so it should not be considered as null risky_weater
+    df['risky_weather'] = np.nan
+    df.loc[df['Temp (°C)'].notna(), 'risky_weather'] = \
+        df['Weather'].astype(str).str.contains('snow|ice|freezing',
+                                   regex=True,
+                                   flags=re.IGNORECASE)
+
     return [(
         int(r['Day']),
         int(r['Time'].split(':')[0]),
@@ -107,7 +119,8 @@ def get_station_weather_month(station_id, year, month):
         float(r['Stn Press (kPa)']),
         float(r['Hmdx']),
         float(r['Wind Chill']),
-        float(r['Temp (°C)'])
+        float(r['Temp (°C)']),
+        float(r['risky_weather']) if not np.isnan(r['risky_weather']) else None
     ) for i, r in df.iterrows()]
 
 
@@ -132,7 +145,8 @@ def get_weather_station_weather_df(spark, stations_id):
             StructField('stn_press', FloatType()),
             StructField('hmdx', FloatType()),
             StructField('wind_chill', FloatType()),
-            StructField('temp', FloatType())
+            StructField('temp', FloatType()),
+            StructField('risky_weather', FloatType())
         ])))
 
     month_per_year_df = spark.createDataFrame(zip(range(1, 13),), ['month'])
@@ -147,6 +161,7 @@ def get_weather_station_weather_df(spark, stations_id):
     create_date_udf = udf(create_date, DateType())
 
     df = (stations_months_df
+          .repartition(200, 'year', 'month')
[…]

### 2019-04-18 preprocess.py
Change street level and street type indexing following The Elements of Statistical Learning section 9.2.4 recommendations

@@ -428,14 +428,8 @@ def get_dataset_df(spark, pos_samples, neg_samples):
 
     pos_samples = pos_samples.select(*neg_samples.columns)
     df = pos_samples.union(neg_samples)
-    street_level_indexer = StringIndexer(inputCol="street_level",
-                                         outputCol="street_level_indexed",
-                                         stringOrderType="alphabetAsc")
-    street_type_indexer = StringIndexer(inputCol="street_type",
-                                        outputCol="street_type_indexed",
-                                        handleInvalid='keep')
-    df = street_level_indexer.fit(df).transform(df).drop('street_level')
-    df = street_type_indexer.fit(df).transform(df).drop('street_type')
+    street_lvl_index, df = index_street_categories(spark, pos_samples, df, 'street_level')
+    street_typ_index, df = index_street_categories(spark, pos_samples, df, 'street_type')
 
     assembler = VectorAssembler(outputCol="features",
                                 inputCols=features_col,
@@ -447,3 +441,29 @@ def get_dataset_df(spark, pos_samples, neg_samples):
                   'date', 'hour', 'features', 'label'))
 
     return df.withColumn('id', monotonically_increasing_id())
+
+
+def index_street_categories(spark, pos_samples, df, col_name): 
+    road = get_road_features_df(spark)
+    street_cat_distrib = (road
+                           .select(col_name)
+                           .na.fill("unknown")
+                           .groupBy(col_name).count()
+                           .withColumn('p', col('count') / lit(road.count()))
+                           .orderBy(col('p').desc())
+                           .drop('count'))
+    street_cat_indexes = \
+        (pos_samples
+         .select(col_name)
+         .na.fill("unknown")
+         .groupBy(col_name).count()
+         .withColumn('p_pos', col('count') / lit(pos_samples.count()))
+         .drop('count')
+         .join(street_cat_distrib, col_name)
+         .na.fill(0, ['p_pos'])
+         .select(col_name, (col('p_pos') - col('p')).alias('risk'))
+         .withColumn(col_name + '_indexed',
+                     row_number().over(Window.orderBy(col('risk').desc())))
+        . drop('risk'))
+
+    return street_cat_indexes, df.join(street_cat_indexes, col_name).drop(col_name)

### 2019-04-18 preprocess.py
Change the weights used for weather information interpolation now use inverse of square of distance

@@ -325,7 +325,7 @@ def get_positive_samples(spark, road_df=None, weather_df=None,
 
 def get_weather_information(samples, weather_df):
     '''Add weather coloumn to samples dataframe. '''
-    p = 1
+    p = 2
     weather_cols = list(set(weather_df.columns)
                         - set(['station_id', 'hour', 'station_lat',
                                'station_long', 'date']))

### 2019-04-18 preprocess.py
Disable least important features

@@ -390,14 +390,14 @@ features_col = ['hour_cos',
                 'street_level_indexed',
                 'street_length',
                 'street_type_indexed',
-                'wind_dir',
+                # 'wind_dir',
                 'rel_hum',
-                'wind_spd',
-                'dew_point_temp',
+                # 'wind_spd',
+                # 'dew_point_temp', corelated with rel_hum and less important
                 'visibility',
                 'stn_press',
-                'wind_chill',
-                'hmdx',
+                # 'wind_chill', mostly nans
+                # 'hmdx', mostly nans
                 'temp',
                 'risky_weather',
                 'dayofyear_cos',
@@ -443,7 +443,7 @@ def get_dataset_df(spark, pos_samples, neg_samples):
     return df.withColumn('id', monotonically_increasing_id())
 
 
-def index_street_categories(spark, pos_samples, df, col_name): 
+def index_street_categories(spark, pos_samples, df, col_name):
     road = get_road_features_df(spark)
     street_cat_distrib = (road
                            .select(col_name)

### 2019-04-18 preprocess.py
Add random forest training notebook

@@ -390,6 +390,7 @@ features_col = ['hour_cos',
                 'street_level_indexed',
                 'street_length',
                 'street_type_indexed',
+                'accident_count',
                 # 'wind_dir',
                 'rel_hum',
                 # 'wind_spd',
@@ -440,7 +441,7 @@ def get_dataset_df(spark, pos_samples, neg_samples):
                   'street_id',
                   'date', 'hour', 'features', 'label'))
 
-    return df.withColumn('id', monotonically_increasing_id())
+    return street_lvl_index, street_typ_index, df.withColumn('id', monotonically_increasing_id())
 
 
 def index_street_categories(spark, pos_samples, df, col_name):

### 2019-04-18 random_undersampler.py
Fix RandomUnderSampler: did not work when the imbalance ratio was already lower than the target imbalance ratio

@@ -58,8 +58,12 @@ class RandomUnderSampler(Estimator, HasTargetImbalanceRatio, HasSeed,
         pos_samples = dataset.filter(col(self.getLabelCol()) == 1.0)
         current_ratio = neg_samples.count()/pos_samples.count()
         sampling = self.getTargetImbalanceRatio()/current_ratio
-        indexes_to_remove = (neg_samples.select(self.getIndexCol())
-                             .sample(1-sampling, seed=self.getSeed()))
+        if sampling > 1.0:
+            # Nothing to do imbalance ratio already lower than target
+            indexes_to_remove = None
+        else:
+            indexes_to_remove = (neg_samples.select(self.getIndexCol())
+                                 .sample(1-sampling, seed=self.getSeed()))
 
         return (RandomUnderSamplerModel(indexes_to_remove, self.getSeed())
                 .setTargetImbalanceRatio(self.getTargetImbalanceRatio()))
@@ -71,6 +75,8 @@ class RandomUnderSamplerModel(Transformer, HasTargetImbalanceRatio):
         self.indexesToRemove = indexesToRemove
 
     def _transform(self, dataset):
+        if self.indexesToRemove is None:
+            return dataset
         index_col = self.indexesToRemove.columns[0]
         return (dataset
                 .join(self.indexesToRemove.withColumn('exists', lit(1.0)),

### 2019-04-18 main_random_forest.py
Change train/test ratio to 0.8/0.2

@@ -13,15 +13,15 @@ from workdir import workdir
 
 spark = init_spark()
 
-neg_samples = get_negative_samples(spark).sample(0.05)
-pos_samples = get_positive_samples(spark).sample(0.05)
+neg_samples = get_negative_samples(spark).sample(0.3)
+pos_samples = get_positive_samples(spark).sample(0.3)
 df = get_dataset_df(spark, pos_samples, neg_samples)
 df_sample = df.filter(col('date') > datetime.fromisoformat('2017-01-01'))
-(train_set, test_set) = df_sample.randomSplit([0.7, 0.3])
+(train_set, test_set) = df_sample.randomSplit([0.8, 0.2])
 
 model = random_forest_tuning(train_set)
 
-with open(workdir + 'data/random_forest_tuning_results_7.txt', 'w') as file:
+with open(workdir + 'data/random_forest_tuning_results_9.txt', 'w') as file:
     for model, result in zip(model.subModels, model.validationMetrics):
         file.write('==================================\n')
         for stage in model.stages:
@@ -32,8 +32,8 @@ with open(workdir + 'data/random_forest_tuning_results_7.txt', 'w') as file:
 
 predictions = model.transform(test_set)
 area_under_PR, f1_score = evaluate_binary_classifier(predictions)
-with open(workdir + 'data/random_forest_tuning_perf_7.txt', 'w') as file:
+with open(workdir + 'data/random_forest_tuning_perf_9.txt', 'w') as file:
     file.write(f"Area Under PR = {area_under_PR}\nF1 score = {f1_score}")
 
 graph = compute_precision_recall_graph(predictions, 20)
-graph.to_csv(workdir + 'data/precision_recall_graph_rf_7.csv')
+graph.to_csv(workdir + 'data/precision_recall_graph_rf_9.csv')

### 2019-04-18 main_xgboost.py
Fix main_xgboost.py and reduce number of parameters tried

@@ -47,7 +47,7 @@ def grid_search(spark):
     model = pipeline.fit(trainDF)
     paramGrid = (ParamGridBuilder().addGrid(xgboost.max_depth, [x for x in range(3, 20, 6)])
                                    .addGrid(xgboost.eta, [x for x in np.linspace(0.2, 0.6, 4)])
-                                   .addGrid(xgboost.scale_pos_weight, [x for x in np.linspace(0.03, 1.0, 12)])
+                                   .addGrid(xgboost.scale_pos_weight, [x for x in np.linspace(0.03, 1.0, 3)])
                                    .build())
     evaluator = BinaryClassificationEvaluator(labelCol="label",
                                               rawPredictionCol="probabilities",
@@ -66,9 +66,8 @@ def grid_search(spark):
                         .asInstanceOf[XGBoostClassificationModel])
 
     with open(workdir + 'data/xgboost_tuning_results_1.txt', 'w') as file:
-        for models, result in zip(model.subModels, model.avgMetrics):
+        for model, result in zip(model.subModels[0], model.avgMetrics):
             file.write('==================================\n')
-            model = models[0]
             for stage in model.stages:
                 params = stage.extractParamMap()
                 for k in params:

### 2019-04-19 utils.py
Fix utils.py: add workdir

@@ -4,6 +4,7 @@ from pyspark import SparkConf
 import requests
 from requests.adapters import HTTPAdapter
 from requests.packages.urllib3.util.retry import Retry
+from workdir import workdir
 
 
 def raise_parquet_not_del_error(cache):
@@ -26,12 +27,14 @@ def init_spark():
                     "org.apache.spark.serializer.KryoSerializer")
             # In local mode there is only one executor: the driver
             # .config("spark.executor.memory", "1536m")
-            .config("spark.driver.memory", "3g")
+            #.config("spark.driver.memory", "3g")
             # Prevent time out errors see: https://github.com/rjagerman/mammoth
             # /wiki/ExecutorLost-Failure:-Heartbeat-timeouts
             #.config("spark.cleaner.periodicGC.interval", "5min")
             #.config("spark.network.timeout", "300s")
-            .config("spark.driver.extraClassPath", "data/xgboost4j-spark-0.72.jar:data/xgboost4j-0.72.jar")
+            .config("spark.driver.extraClassPath",
+                    f"{workdir}data/xgboost4j-spark-0.72.jar"
+                    + f":{workdir}data/xgboost4j-0.72.jar")
             .getOrCreate())
 
     print('Spark Session created')

### 2019-04-19 main_brf.py
Add balanced random forest tuning

@@ -1,0 +1,42 @@
+#!/usr/bin/env python3
+from datetime import datetime
+from pyspark.sql.functions import col
+from random_forest import balanced_random_forest_tuning, \
+                          compute_precision_recall, \
+                          compute_precision_recall_graph
+from preprocess import get_positive_samples, \
+                       get_negative_samples, \
+                       get_dataset_df
+from evaluate import evaluate_binary_classifier
+from utils import init_spark
+from workdir import workdir
+
+spark = init_spark()
+
+i=1
+sampleFraction=0.01
+
+neg_samples = get_negative_samples(spark).sample(sampleFraction)
+pos_samples = get_positive_samples(spark).sample(sampleFraction)
+df = get_dataset_df(spark, pos_samples, neg_samples)
+(train_set, test_set) = df.randomSplit([0.8, 0.2])
+(train_set, test_set) = (train_set.persist(), test_set.persist())
+
+model = balanced_random_forest_tuning(train_set)
+
+with open(workdir + f'data/brf_tuning_results_{i}.txt', 'w') as file:
+    for model, result in zip(model.subModels[0], model.avgMetrics):
+        file.write('==================================\n')
+        for stage in model.stages:
+            params = stage.extractParamMap()
+            for k in params:
+                file.write(f'{k.name}: {params[k]}\n')
+        file.write(f"Area under PR: {result}\n")
+
+predictions = model.transform(test_set)
+area_under_PR, f1_score = evaluate_binary_classifier(predictions)
+with open(workdir + f'data/brf_tuning_perf_{i}.txt', 'w') as file:
+    file.write(f"Area Under PR = {area_under_PR}\nF1 score = {f1_score}")
+
+graph = compute_precision_recall_graph(predictions, 20)
+graph.to_csv(workdir + f'data/precision_recall_graph_brf_{i}.csv')

### 2019-04-19 main_rf_test.py

@@ -1,37 +1,0 @@
-#!/usr/bin/env python3
-from datetime import datetime
-from pyspark.ml.classification import RandomForestClassifier
-from pyspark.sql.functions import col
-from preprocess import get_positive_samples, \
-                       get_negative_samples, \
-                       get_dataset_df, \
-                       random_undersampling
-from utils import init_spark
-from workdir import workdir
-from random_forest import compute_precision_recall_graph
-from evaluate import evaluate_binary_classifier
-
-spark = init_spark()
-
-neg_samples = get_negative_samples(spark)
-pos_samples = get_positive_samples(spark)
-
-neg_samples = random_undersampling(pos_samples, neg_samples, target_ratio=2.0)
-
-df = get_dataset_df(spark, pos_samples, neg_samples)
-df_sample = (df
-             .filter(col('date') > datetime.fromisoformat('2017-01-01'))
-             .persist())
-
-(train_set, test_set) = df_sample.randomSplit([0.7, 0.3])
-
-rf = RandomForestClassifier(labelCol="label",
-                            featuresCol="features",
-                            cacheNodeIds=True,
-                            maxDepth=30)
-model = rf.fit(train_set)
-predictions = model.transform(test_set)
-evaluate_binary_classifier(predictions)
-graph = compute_precision_recall_graph(predictions, 20)
-graph.to_csv(workdir + 'data/precision_recall_graph_rf_'
-             + 'randomundersampling2.csv')

### 2019-04-19 random_forest.py

@@ -1,7 +1,9 @@
 from pyspark.ml.classification import RandomForestClassifier
 from pyspark.ml.evaluation import BinaryClassificationEvaluator
 from pyspark.ml.evaluation import MulticlassClassificationEvaluator
-from pyspark.ml.tuning import TrainValidationSplit, ParamGridBuilder
+from pyspark.ml.tuning import TrainValidationSplit, \
+                              ParamGridBuilder, \
+                              CrossValidator
 from pyspark.ml import Pipeline
 from pyspark.sql.types import DoubleType
 from pyspark.sql.functions import udf, col
@@ -9,6 +11,7 @@ import pandas as pd
 import numpy as np
 from preprocess import features_col
 from random_undersampler import RandomUnderSampler 
+from class_weighter import ClassWeighter
 
 
 def random_forest_tuning(train_samples):
@@ -22,9 +25,9 @@ def random_forest_tuning(train_samples):
          .addGrid(rf.numTrees, [50, 75, 100])
          .addGrid(rf.featureSubsetStrategy, ['sqrt'])
          .addGrid(rf.impurity, ['gini', 'entropy'])
-         .addGrid(rf.maxDepth, [30])
+         .addGrid(rf.maxDepth, [5, 15, 30])
          .addGrid(rf.minInstancesPerNode, [1])
-         .addGrid(rf.subsamplingRate, [0.6, 0.5, 0.4])
+         .addGrid(rf.subsamplingRate, [1.0, 0.6, 0.4])
          .addGrid(ru.targetImbalanceRatio, [1.0, 1.5, 2.0])
          .build())
     pr_evaluator = \
@@ -34,12 +37,47 @@ def random_forest_tuning(train_samples):
     tvs = TrainValidationSplit(estimator=pipeline,
                          estimatorParamMaps=paramGrid,
                          evaluator=pr_evaluator,
-                         trainRatio=0.7,
+                         trainRatio=0.8,
                          collectSubModels=True)
 
     model = tvs.fit(train_samples)
 
     return model
+
+
+def balanced_random_forest_tuning(train_samples):
+    rf = RandomForestClassifier(labelCol="label",
+                                featuresCol="features",
+                                cacheNodeIds=True,
+                                weightCol="weight")
+    ru = RandomUnderSampler().setIndexCol('id')
+    cw = ClassWeighter()
+    pipeline = Pipeline().setStages([ru, cw, rf])
+    paramGrid = \
[…]

### 2019-04-19 paper/paper.tex
Change positive and negative examples generation part of paper

@@
 Road accidents are an important issue of our modern societies, responsible for millions of deaths and injuries.
-In this paper we show how one can leverage open datasets from a City like Montreal, Canada, to create some accident prediction models, using state-of-the-art big data analytics methods in Python.
+In this paper we show how one can leverage open datasets from a city like Montreal, Canada, to create some accident prediction models, using state-of-the-art big data analytics methods in Python.
 Such models could then be used in the context of road accident prevention, but also to identify key factors that can lead to a road accident, and eventually, help to elaborate new policies.
@@
 All source code can be found on Github.
+
 \end{abstract}
@@
 In the context of accident prediction, Big Data Analytics could provide insights on the conditions leading to an increased risk of road accidents that could then be used to develop traffic-related policies and prevention operations.
-Our contributions include: (1) a demonstration of how simple and easily available datasets can be assembled to obtain meaningful features for road accident prediction, (2) a comparison of three different machine learning algorithm dealing with data imbalance in the context of road accident prediction, (3) the implementation of Balanced Random Forest introduced by Chen et al.
-in Apache Spark for efficient distributed training.
-Finally, all the source code used is publicly available on Github under the MIT license.
+Our contributions include:
+\begin{itemize}
+\item a demonstration of how simple and easily available datasets can be assembled to obtain meaningful features for road accident prediction,
+\item a comparison of three different machine learning algorithm dealing with data imbalance in the context of road accident prediction,
+\item the implementation of Balanced Random Forest introduced by Chen et al. in Apache Spark for efficient distributed training. 
+\end{itemize}
+All the source code used is publicly available on Github under the MIT license.\footnote{https://github.com/GTimothee/accident-prediction-montreal}
 
@@
 \cite{Chang2005, Theofilatos2017, Abellan2013, Lin2015}, along with other machine learning algorithms like Support Vector Machines \cite{LI20081611}.
-Recently, the outstanding progress of deep learning algorithms have resulted in their extensive use in every field, including in Geospatial machine learning (Convolutional Neural Network and Long Short-Term Memory neural network for example) \cite{Yuan2018, QChen2016}.
+Recently, the outstanding progress of deep learning algorithms has resulted in their extensive use in every field, including in Geospatial machine learning (Convolutional Neural Network and Long Short-Term Memory neural network for example) \cite{Yuan2018, QChen2016}.
 \section{Methods}
@@
 \subsection{Datasets}
-We used three public datasets provided by the city of Montreal and the government of Canada: (1) Montreal Vehicle Collision: This dataset provided by the city of Montreal contains all the road collisions reported by the police in Montreal from 2012 to 2018.
+We used three public datasets provided by the city of Montreal and the government of Canada: 
+\begin{itemize}
+\item Montreal Vehicle Collision: This dataset provided by the city of Montreal contains all the road collisions reported by the police in Montreal from 2012 to 2018.
 For each accident, the dataset contains the date and localization of the accident, information on the number of injuries and death, the number of vehicles involved, and information on the road condition.
@@
 Another dataset with all vehicle collision in Canada is available but without the localization of the accident, therefore we restrained our analysis to the city of Montreal.
-(2) National Road Network: This dataset provided by the government of Canada contains the geometry of all roads in Canada.
+\item National Road Network: This dataset provided by the government of Canada contains the geometry of all roads in Canada.
 For each road segment, a few meta-data are given.
@@
 The data was available in various format, we chose to use the Keyhole Markup Language, which is easy to read because it is based on the Extensible Markup Language (XML).
-(3) Data Historical Climate Data: This dataset provided by the government of Canada contains hourly weather information measured at different weather stations across Canada.
+\item Data Historical Climate Data: This dataset provided by the government of Canada contains hourly weather information measured at different weather stations across Canada.
 For each station and every hour, the dataset provides the temperature, the humidity, the wind direction and speed, the visibility, the atmospheric pressure and observations of atmospheric phenomenon such as snow, fog, rain, etc.
+\end{itemize}
 
-\subsection{Positive and negative samples generation}
-Our machine learning problem can be stated as a binary classification problem, the positive class being the occurrence of an accident and the negative class being all the cases where an accident did not happen.
-We got a little bit less than 150 000 positive samples from the national collision dataset and we generated a random sub-sample of 5 million negative samples (over 10 billion possibilities).
-The generation of the negative samples basically consisted in a Cartesian product between the roads from the road network dataset and all the hours of all days that occurred between the 1st January 2012 and the 31 of December 2017, the dates range of the Montreal Vehicle Collision dataset.
-This huge cartesian product operation together with the joins with the data from the other datasets made our dataset generation pretty heavy.
+\subsection{Positive and negative examples generation}
+Our machine learning problem can be stated as a binary classification problem, the positive class being the occurrence of an accident and the negative class being the non-occurence of an accident on a given road at a given date and hour.
+The vehicle colission dataset contains 150 thousands of collisions, among them $134 489$ contain the date, the hour and the localisation of the accident.
+For each of these accidents, we identify the closest road segment.
+These pairs of time and  road segment are used as positive examples.
+For the negative examples, we generated a random sample of 5 millions combinations of time and road segment which did not appear in the collision dataset.
+The generation of these examples together with the join operation with the data from the other datasets made our dataset generation pretty heavy.
 It resulted in an impractical amount of time and memory space requirements to generate the dataset.
@@
 Finally, the identification of the road segment corresponding to accidents was very memory intensive, we modified this step to be executed by batch of one month.
-With these improvements and a few other tricks including partitioning the data frame at key points in our algorithm, we manage to reduce the processing time from about one minute to generate about 1000 positive samples to less than 1 second.
+With these improvements and a few other tricks including partitioning the data frame at key points in our algorithm, we managed to reduce the processing time to a reasomable time.
 We also used a cluster from “Compute Canada” to take maximum advantage of Apache Spark distributed nature for the generation of examples and the hyper-parameter tuning of our models.
@@
 In addition road segment are classified in three different level in the dataset depending on their importance in the road network, we created a categorical feature from this information.
-For these two categorical features, we encoded them as suggested in Element of Statistical Learning\cite{elementsofstat} in section 9.2.4, instead of using one hot encoding which would create an exponential number of splits, we index the categorical variable ordered by the proportion of the example belonging to the given category which are positive samples.
+For these two categorical features, we encoded them as suggested in The Elements of Statistical Learning\cite{elementsofstat} in section 9.2.4, instead of using one hot encoding which would create an exponential number of splits, we index the categorical variable ordered by the proportion of the example belonging to the given category which are positive samples.
 This encoding guarantee to provide optimal splits on these categorical variables.
@@
 
- Our source code is available on Github at \href{https://github.com/GTimothee/accident-prediction-montreal}{https://github.com/GTimothee/accident-prediction-montreal}.
-
 \section{Discussion}

### 2019-04-19 paper/paper.tex
Correct language mistakes

@@
 In particular, we show how we implemented Balanced Random Forests, a variant of the Random Forests machine learning algorithm, into the Scala and Python APIs of Apache Spark big data framework.
-All source code can be found on Github.
 
@@
 \item a comparison of three different machine learning algorithm dealing with data imbalance in the context of road accident prediction,
-\item the implementation of Balanced Random Forest introduced by Chen et al. in Apache Spark for efficient distributed training. 
+\item the implementation of Balanced Random Forest introduced by Chen et al.\cite{Chen2004} in Apache Spark for efficient distributed training. 
 \end{itemize}
@@
 However, we found ourselves writing a lot more code using Dask than Spark for the same result.
-Spark map-reduce-based interface is very handy and allow easier processing of datasets.
+Spark map-reduce-based interface is very handy and allows easier processing of datasets.
 For all these reasons we found out that Spark fitted our needs best and we switched to it at an early time of the study.
@@
 In particular, we implemented the Balanced Random Forests (BRF) algorithm in Apache Spark to deal with class imbalance which is a prominent issue in road accident prediction.
-To validate the performance of our version, we run an experiment with a simple synthetic dataset to compare its performances with the standard implementation of Random Forest (RF) in Spark with random under-sampling of the majority class.
-Balanced Random Forest performed much better.
-Balanced Random Forests in one of the two methods proposed by Chen, Liaw, and Breiman to deal with class imbalance when using Random Forest.
+To validate the performance of our version, we run an experiment with a simple synthetic dataset to compare its performances with the standard implementation of Random Forest (RF) in Spark using random under-sampling of the majority class.
+As expected, Balanced Random Forest performed better.
+Balanced Random Forests in one of the two methods proposed by Chen, Liaw, and Breiman\cite{Chen2004} to deal with class imbalance when using Random Forest.
 Balanced Random Forests are like Random Forests, but with a difference during the bootstrapping phase, for each tree of the forest, a random under-sampling of the majority class is performed in order to obtain a balanced sample.
@@
 Chen, Liaw and Breiman\cite{Chen2004} also proposed in the same paper, the Weighted Random Forest (WRF) which consists in giving more weight to the minority class at the building time of each tree and during the measure of the gain in impurity and during the class prediction of each terminal node.
-While neither method is clearly better than the other in terms of prediction power, we chose BRF because it is found to be more computationally efficient (because of the under-sampling) and easier to implement in our opinion.
+While neither method is clearly better than the other in terms of prediction power, we chose BRF because it is found to be more computationally efficient (because of the under-sampling).
 Interestingly, Wallace et al.
@@
 For each road segment, a few meta-data are given.
-Sadly, for roads in Quebec the speed limit and the surface type are not provided.
-The data was available in various format, we chose to use the Keyhole Markup Language, which is easy to read because it is based on the Extensible Markup Language (XML).
+For roads in Quebec, the speed limit and the surface type are not provided.
+The data was available in various format, we chose to use the Keyhole Markup Language, which is a standard of the Open Geospatial Consortium since 2008.
+This format is based on the Extensible Markup Language (XML), which make it easier to read using existing implementation of XML parsers.
 \item Data Historical Climate Data: This dataset provided by the government of Canada contains hourly weather information measured at different weather stations across Canada.
@@
 \subsection{Positive and negative examples generation}
-Our machine learning problem can be stated as a binary classification problem, the positive class being the occurrence of an accident and the negative class being the non-occurence of an accident on a given road at a given date and hour.
-The vehicle colission dataset contains 150 thousands of collisions, among them $134 489$ contain the date, the hour and the localisation of the accident.
+This machine learning problem can be stated as a binary classification problem, the positive class is the occurrence of an accident and the negative class is the non-occurrence of an accident on a given road at a given date and hour.
+The vehicle collision dataset contains 150 thousands of collisions, among them $134 489$ contain the date, the hour and the localization of the accident.
 For each of these accidents, we identify the closest road segment.
-These pairs of time and  road segment are used as positive examples.
+These pairs of a time and a road segment are used as positive examples.
 For the negative examples, we generated a random sample of 5 millions combinations of time and road segment which did not appear in the collision dataset.
-The generation of these examples together with the join operation with the data from the other datasets made our dataset generation pretty heavy.
+The generation of these examples together with the join operation with the data from the other datasets made our dataset generation heavy.
 It resulted in an impractical amount of time and memory space requirements to generate the dataset.
@@
 Finally, the identification of the road segment corresponding to accidents was very memory intensive, we modified this step to be executed by batch of one month.
-With these improvements and a few other tricks including partitioning the data frame at key points in our algorithm, we managed to reduce the processing time to a reasomable time.
+With these improvements and a few other tricks including partitioning the data frame at key points in our algorithm, we managed to reduce the processing time to a reasonable time.
 We also used a cluster from “Compute Canada” to take maximum advantage of Apache Spark distributed nature for the generation of examples and the hyper-parameter tuning of our models.
@@
 \subsection{Feature Engineering}
-Once we had our pairs of date with hour and road segment for positive and negative examples, we created features from this information.
-We created four types of features, weather features, features from the road segment, features from the date .
+Once we had our pairs of an hour and a road segment for positive and negative examples, we created features from this information.
+We created three types of features, weather features, features from the road segment and features from the date and time.
 
 For the weather features, we used data from the Historical Climate Dataset.
-In order to estimate the weather information at the position of the road segment we compute the mean of the weather information from all  the surrounding weather stations at the date and hour of the example  weighted by the inverse of the square of the distance between the station and the road segment.
+In order to estimate the weather information at the position of the road segment, we use the mean of the weather information from all the surrounding weather stations at the date and hour of the example weighted by the inverse of the square of the distance between the station and the road segment.
 We initially used the inverse of the distance, but we obtained a small improvement in performances when squaring the inverse of the distance.
-It makes sense since, the weather information is probably much closer to the closest stations than the other ones.
-We tried higher power, but results were not as good.
+It makes sense since the weather information is probably much closer to the closest stations than the other ones.
+We tried higher power, but the results were not as good.
 It could be interesting to experiment with different power values for each weather information.
 We used all the weather information provided by the Historical Climate Dataset.
-The features are: the temperature, the dew point temperature which is a measure of the humidity, the real humidty percentage, the wind direction, the wind speed, the visibility, the atmospheric pressure, the Hmdx index, which is a measure of felt temperature and the wind chill which is another measure of the felt temperature using the wind information.
+The features are: the temperature, the dew point temperature which is a measure of the humidity, the real humidity percentage, the wind direction, the wind speed, the visibility, the atmospheric pressure, the Hmdx index, which is a measure of felt temperature and the wind chill which is another measure of the felt temperature using the wind information.
 In addition, the historical weather dataset, provide a multi-label categorical variable providing the observations of atmospheric phenomenon such as snow, fog, rain, etc.
-We used this last variable to create a binary variable that we called risky weather which is true when the following phenomenon are observed: Freezing Rain, Freezing Drizzle, Snow, Snow Grains, Ice Crystals, Ice Pellets, Ice Pellet Showers, Snow Showers, Snow Pellets, Ice Fog, Blowing Snow, Freezing Fog.
+We used this last variable to create a binary variable that we called risky weather which is true when the following phenomenons are observed: Freezing Rain, Freezing Drizzle, Snow, Snow Grains, Ice Crystals, Ice Pellets, Ice Pellet Showers, Snow Showers, Snow Pellets, Ice Fog, Blowing Snow, Freezing Fog.
 
 For the features from the road segments, we were limited by the limited metadata provided on the road segments.
-From the shape of the road segment, we computed the length of the road segment, and from the name of the street we identified the type of road (Highway, street, boulevard, etc.).
-In addition road segment are classified in three different level in the dataset depending on their importance in the road network, we created a categorical feature from this information.
-For these two categorical features, we encoded them as suggested in The Elements of Statistical Learning\cite{elementsofstat} in section 9.2.4, instead of using one hot encoding which would create an exponential number of splits, we index the categorical variable ordered by the proportion of the example belonging to the given category which are positive samples.
+From the shape of the road segment, we computed the length of the road segment, and from the name of the street, we identified the type of road (Highway, street, boulevard, etc.).
+In addition, road segments are classified into three different levels in the dataset depending on their importance in the road network, we created a categorical feature from this information.
+For these two categorical features, we encoded them as suggested in The Elements of Statistical Learning\cite{elementsofstat} in section 9.2.4, instead of using one hot encoding which would create an exponential number of splits, we index the categorical variable ordered by the proportion of the examples belonging to the given category which are positive samples.
 This encoding guarantee to provide optimal splits on these categorical variables.
-Lastly, we added an a feature giving the number of accidents which occurred previously on this road segment.
+Lastly, we added a feature giving the number of accidents which occurred previously on this road segment.
 
-For the date features, is is quite simple, we took the day of the year, the hour of the day, and the day of the week.
-We decided to encode day of the year and hour of the day to be cylic.
-Indeed, for example with hour of the day, 23 hour is very close to 1 hour, but in the usual encoding this does not appear to the machine learning algorithm.
+For the date features, we took the day of the year, the hour of the day, and the day of the week.
+We decided to make the features day of the year and hour of the day cyclic.
+Indeed, for example with the feature hour of the day, 23 hour is very close to 1 hour, but in the usual encoding, this does not appear to the machine learning algorithm.
 With cyclical encoding, we compute two features, the first one is the cosine of the original feature scaled between 0 and two pi, and the second one is the sine of the original feature scaled between 0 and two pi.
+In this form, the hours 23 and 1 appear close.
 
@@
 The Balanced Random Forests algorithm was not implemented in Apache Spark.
-An implementation is available in the python library imbalanced-learn\cite{imbalance} which implements many algorithm to deal with data imbalance using an API inspired by scikit-learn, but the size of our dataset made it impossible for us to use this library.
+An implementation is available in the python library imbalanced-learn\cite{imbalance} which implements many algorithms to deal with data imbalance using an API inspired by scikit-learn, but the size of our dataset made it impossible for us to use this library.
 Therefore, we implemented Balance Random Forests in Apache Spark.
@@
 When doing sampling with replacement, each value of this array is given by a Poisson distribution.
-The parameter of the Poisson distribution correspond to the subsampling rate hyper-parameter of the Random Forest, which specifies the size of the sample used for training each tree as a fraction of the total size of the dataset.
+The parameter of the Poisson distribution corresponds to the subsampling rate hyper-parameter of the Random Forest, which specifies the size of the sample used for training each tree as a fraction of the total size of the dataset.
 Indeed, if for example we want each tree to use a sample of the same size of the whole dataset, the subsampling ratio will be set to 1.0, which is indeed the number of time a given example will appear in a tree on average.
@@
 In order to implement Balanced Random Forests, we modified the parameter of the Poisson distribution to use the class weight multiplied by the subsampling ratio, this way a negative sample with for example the weight 0.25 has 4 times less chance to be chosen to appear in a given tree.
-This implementation has the advantage that it did not required a big code change and will be easy to test, but it also has a disadvantage, users probably expect linearly correlated weight to be equivalent, which is not the case in our implementation.
+This implementation has the advantage that it did not require a big code change and will be easy to test, but it also has a disadvantage, users probably expect linearly correlated weight to be equivalent, which is not the case in our implementation.
 
-In order to be compatible with other possible use cases, the weight are actually applied per samples and not per class.
+In order to be compatible with other possible use cases, the weights are actually applied per samples and not per class.
 This is a choice made by Apache Spark developers that we respected.
@@
 Finally, we did a first validation of our implementation using an artificially generated imbalanced dataset and compared the results of both standard random forest and BRF.
-While BRF had better results in terms of area under PR (which was our selected measure) (98.\% against 98.1\%), standard random forest had better F1 score (99.7\% against 98.8\%) with default threshold.
+While BRF had better results in terms of area under PR (which was our selected measure) (98.\% against 98.1\%), the standard random forest had better F1 score (99.7\% against 98.8\%) with default threshold.
 This emphasized that we were right in our choice of the area under PR curve as our measure in the case of class imbalance, indeed with a good threshold, the BRF model performs better.
@@
 \subsection{Hyper-parameter tunings}
-In order to determine the optimal hyper parameter, we first performed automatic hyper parameter tuning using Apache Spark ML library grid search using cross validation.
+In order to determine the optimal hyperparameter, we first performed automatic hyperparameter tuning using Apache Spark ML library grid search using cross-validation.
 Because the training on the whole dataset would have been too high, we took a small sample of the dataset.
@@
 
-Once we got a first tuning result with grid search we continued manually by following a plan, do , check, adjust method, using the measure of the area under the precision recall curve, the precision and recall with different threshold on the test set and the training set we were able to better understand how the performances of our model could be improved.
+Once we got a first tuning result with grid search we continued manually by following a plan, do , check, adjust method, using the measure of the area under the precision-recall curve, the precision and recall with different threshold on the test set and the training set we were able to better understand how the performances of our model could be improved.
 
 Interestingly, despite using many trees, our Random Forest algorithms tended to overfit very quickly as soon as the maximum depth parameter goes above 18.
-We eventually used only 100 trees, because more trees did not help much.
-We have not tried more than 200 trees, maybe many more trees would have been necessary to increase the maximum depth without over fitting, but then the training time would not be reasonable.
+We eventually used only 100 trees, because adding more trees did not increase performances.
+We have not tried more than 200 trees, maybe many more trees would have been necessary to increase the maximum depth without overfitting, but then the training time would not be reasonable.
 
 \subsection{Choice of a metric to evaluate our models}
-The Chosen metric for the evaluation of the models was the area under Precision/Recall (PR) curve (area under PR).
+The chosen metric for the evaluation of the models was the area under the precision-recall curve (area under PR).
 According to Davis and Goadrich\cite{Davis} this measure is the most representative of how well a model has been trained in the case of data imbalance, particularly compared to the ROC curve which is the commonly used measure.
@@
 According to this figure, the most important features to predict the occurrence of an accident is the number of accidents which occurred on this road during the previous years which we could have easily guessed.
-Then the location of the road, the type and length of the road segment and the hour of the day have a high importance.
+Then the location of the road, the type and length of the road segment, and the hour of the day have high importance.
 As compared to the count of accident feature the remaining features seem to have almost no importance, but the model does not perform as well if we remove them.
 Surprisingly, the day of the year, the temperature and the risky weather boolean have very low feature importance.
-We believe that the weather features have lower feature importances because they are strongly correlated.
+We believe that the weather features have lower importances because they are strongly correlated.
 The gain in impurity the information they contain provide is distributed over all the correlated features.
@@
 
-These results have been obtained by training the Balanced Random Forest algorithm on the whole dataset of positive samples, but with only a fifth of the dataset of negative samples we have generated.
-Therefore the data imabalance is reduced to a factor of 8.
-According to our previous experiments, the results are almost the same when using the full dataset of negative samples, sometimes slightly better.
+These results have been obtained by training the Balanced Random Forest algorithm on the whole dataset of positive samples, but with only a fifth of the dataset of negative examples we have generated.
+Therefore the data imbalance is reduced to a factor of 8.
+According to our previous experiments, the results are almost the same when using the full dataset of negative examples, sometimes slightly better.
 To evaluate this model we used a test set containing the last two years of our dataset, the model is trained on the 4 previous years and use only data from these years.
@@
 
-We obtain an area under the precision-recall curve of 79\%, and an area under the ROC curve of 97\%.
+We obtain an area under the precision-recall curve of 79\% and an area under the ROC curve of 97\%.
 This results in the following precision and recall values depending on the chosen probability threshold:
@@
 As we can see, with a threshold set to 0.63, we can achieve a recall of 85\% with a precision of 0.52\%.
-With a threshold of 0.06, the model can actually detect almost all the accidents, 94\% of them, with a precision of 37\%, this corrsponds to the high-risk roads and time in which we predict a risk of 37\% of accidents.
+With a threshold of 0.06, the model can actually detect almost all the accidents, 94\% of them, with a precision of 37\%, this corresponds to the high-risk roads and time in which we predict a risk of 37\% of accidents.
 
 \section{Discussion}
-The overall results are quite good but mostly rely on the count of previous accidents on the road segment.
+The overall results are good but mostly rely on the count of previous accidents on the road segment.
 This is not an issue for accident prediction, but it does not help to understand why these roads are particularly dangerous.
@@
 Therefore we believe providing more context information on where the road is located will greatly help.
-On another experiment we have seen the population density and the distance to the closest billboard used as features successfully.
+In another experiment, we have seen the population density and the distance to the closest billboard used as features successfully.
 However, it will be hard to reach good performances because of the nature of the problem without the count of accidents.
@@
 These datasets might not be as easily available for over areas, but for Montreal, it could improve performances.
-Also, the solar elevation and the solar azimuth could easily be added without the need for another dataset.
+Also, the solar elevation and solar azimuth could easily be added without the need for another dataset.
 These features have been used successfully in road accident prediction in previous experiments.

### 2019-04-20 road_network.py
Fix negative sample generation: filter out road segment outside of the zone of the collision dataset

@@ -7,12 +7,16 @@ from zipfile import ZipFile
 from shutil import rmtree
 from io import BytesIO
 from bs4 import BeautifulSoup
+from pyspark.sql import Window
 from pyspark.sql.functions import col, abs, hash, atan2, \
                                   sqrt, cos, sin, radians, \
-                                  udf, monotonically_increasing_id
+                                  udf, monotonically_increasing_id, \
+                                  concat, row_number
 from pyspark.sql.types import StringType
 from utils import raise_parquet_not_del_error
 from workdir import workdir
+from road_network_nids import unknow_file_included_nids
+
 
 
 def get_road_df(spark, use_cache=True):
@@ -79,18 +83,38 @@ def fetch_road_network():
         print('Skip fetching road network: already downloaded')
         return
     print('Fetching road network...')
-    try:
-        url = 'http://ftp.maps.canada.ca/pub/nrcan_rncan'\
+    url = 'http://ftp.maps.canada.ca/pub/nrcan_rncan'\
                 '/vector/geobase_nrn_rrn/qc/kml_en/'
-        html_list = urlopen(url)
-        bs = BeautifulSoup(html_list, 'lxml')
-    except (URLError, HTTPError):
-        print('Unable to reach dataset server')
-        raise
-    files = map(lambda tr: tr.a.text, bs.body.table.find_all('tr')[3:-1])
-    files = filter(lambda f: re.match('.*_[4-7]_(5[5-9]|60).kmz', f), files)
     if not os.path.isdir(workdir + 'data/road-network'):
         os.mkdir(workdir + 'data/road-network')
+    files = \
+        ['TNO_terrestre_du_TE_de_Montréal_6_60.kmz', 'Mont-Royal_4_58.kmz',
+         "L'Île-Dorval_4_57.kmz", 'Montréal_4_58.kmz', 'Unknown_5_57.kmz',
+         'Unknown_5_58.kmz', 'Unknown_4_55.kmz', 'Montréal_5_59.kmz',
+         'Unknown_5_56.kmz', 'Montréal_4_55.kmz', "L'Île-Dorval_4_58.kmz",
+         'Mont-Royal_5_59.kmz', 'Montréal_4_60.kmz', 'Montréal_4_56.kmz',
+         'Montréal_6_60.kmz', 'Montréal-Ouest_4_58.kmz',
+         'Senneville_4_56.kmz', 'Montréal_6_58.kmz', 'Unknown_4_59.kmz',
[…]

### 2019-04-20 road_network_nids.py

@@ -1,0 +1,1774 @@
+# List obtained by graphically selection on google earth the road inside the island of Montreal
+unknow_file_included_nids = ['73671c5348054b448b94358ac26ff548',
+'0de4c13b13334812ba1803a4d2ef1cf8',
+'72d5c5f727614966b8a3733c0a2c032d',
+'c34dadb8dcf24143b3d14eab75dae974',
+'72beb531ce7f476fa383616a7ed7276b',
+'3f26f14865b244a9acf31f812dc83aa0',
+'b26c24b5349442049a74742dd4c9a3f9',
+'6b8c930b1f53480a9eca6365c3c4a7e7',
+'faf1cb7e937d44639440e1eefff5adea',
+'8ec66e8fa96643618cefe8ed8735eef6',
+'e3fa30f1cd8641beab219f00a32d391d',
+'15a3d60020824b2285030b7fa77f0a6c',
+'3c84b65d022a401190a0de49858a5e22',
+'3408fd326abe43af86a61250b1111efa',
+'40086095074e446b96a36865de7a8679',
+'e6c2a08cf30c4566afdfd2832080d5d3',
+'ec0d0ad9a5334d4ca5893638969863f1',
+'d6695b4cc4cf46189eac7561d1fa72a2',
+'b5b3fe8dd24348ad99fcdd0282efc7ab',
+'faf36224560f4f82b18aa08a7b889c8c',
+'4bbcd933bcaf42a3a95b1a69efefb8b4',
+'a67f895a006c425289a84a0225fd8b96',
+'fc0225d59e43435488ea1f814735b6ad',
+'30cd4380cc454645b1a8cb85c0a240b9',
+'d8d16706c88a485884b5284324167076',
+'73d4bf983e8647cc880441e41ea3bc8c',
+'aa0294bb10734b4cb1b91392a7a5402c',
+'0787e804523248bfa73ea28e76b5be96',
+'500f55d39dfe4aeb8465d338974dd78c',
+'e64ba475514d4f48ab8ce2ea044cece4',
+'9721d22f9b8f4d0188bcd3220aa05fe7',
+'51ec453db2104bad933c8cf893a12fba',
+'658fabb5d80a44ad9bc9b0f5a01f9aab',
+'54240bcb269b4287bfe04e92356019ec',
+'2d659954e2ba462f9c6be435398bc69c',
[…]

### 2019-04-20 weather.py
Improve risky_weather feature: moving weighted average to express effect of snow in later hours

@@ -2,8 +2,11 @@ import re
 import datetime
 from io import StringIO
 from os.path import isdir
+from math import exp
 from bs4 import BeautifulSoup
-from pyspark.sql.functions import udf, col, year, month, dayofmonth, explode
+from pyspark.sql import Window
+from pyspark.sql.functions import udf, col, year, month, dayofmonth, \
+                                  explode, lag, coalesce, lit
 from pyspark.sql.types import StructField, FloatType, StructType, \
                               IntegerType, ArrayType, DateType
 from requests import get
@@ -179,6 +182,30 @@ def get_weather_station_weather_df(spark, stations_id):
                   c['wind_chill'].alias('wind_chill'),
                   c['temp'].alias('temp'),
                   c['risky_weather'].alias('risky_weather')))
+
+    # We make a moving average of risky_weather since the effect of a risky
+    # weather is distributed in the next hours
+    def weighted_average(c, window, offsets, weights):
+        def value(i):
+            return lag(c, -i).over(window)
+
+        values = [coalesce(value(i) * w, lit(0)) 
+                for i, w in zip(offsets, weights)]
+
+        return sum(values, lit(0))
+
+    window = (Window
+            .partitionBy('station_id')
+            .orderBy('date'))
+    offsets = range(-23, 1)
+    l = 0.5
+    weights = [exp(l*t) for t in offsets]
+    weights = [w/sum(weights) for w in weights]
+    df = df.withColumn('risky_weather',
+                            weighted_average('risky_weather',
+                                            window,
+                                            offsets,
+                                            weights))
 
     df.write.parquet(cache_file)
 

### 2019-04-20 paper/paper.tex
Update results with correction of negative examples with 0.5 sample of 1/1000 dataset

@@
 These pairs of a time and a road segment are used as positive examples.
-For the negative examples, we generated a random sample of 5 millions combinations of time and road segment which did not appear in the collision dataset.
+For the negative examples, we generated a random sample of 2 millions combinations of time and road segment which did not appear in the collision dataset.
 The generation of these examples together with the join operation with the data from the other datasets made our dataset generation heavy.
@@
 \section{Results}
-The training time of the machine learning algorithm is very long due to the size of our dataset and the latest improvement in term of feature engineering have been discovered very late.
-Therefore, we currently only have the results with the Balanced Random Forest using the latest features.
-The results from the previous training of other algorithm are not comparable since they do not use the same features.
-We will add the missing results in later versions of this document.
-
-These results have been obtained by training the Balanced Random Forest algorithm on the whole dataset of positive samples, but with only a fifth of the dataset of negative examples we have generated.
+These results have been obtained by training the Balanced Random Forest algorithm on the whole dataset of positive samples, but with only half of the dataset of negative examples we have generated.
 Therefore the data imbalance is reduced to a factor of 8.
-According to our previous experiments, the results are almost the same when using the full dataset of negative examples, sometimes slightly better.
 To evaluate this model we used a test set containing the last two years of our dataset, the model is trained on the 4 previous years and use only data from these years.
@@
 
-We obtain an area under the precision-recall curve of 79\% and an area under the ROC curve of 97\%.
+We obtain an area under the precision-recall curve of 54\% and an area under the ROC curve of 89\%.
 This results in the following precision and recall values depending on the chosen probability threshold:
@@
 
-As we can see, with a threshold set to 0.63, we can achieve a recall of 85\% with a precision of 0.52\%.
-With a threshold of 0.06, the model can actually detect almost all the accidents, 94\% of them, with a precision of 37\%, this corresponds to the high-risk roads and time in which we predict a risk of 37\% of accidents.
+As we can see, with a threshold set to 0.63, we can achieve a recall of 85\% with a precision of 0.29\%.
+With a threshold of 0.05, the model can actually detect almost all the accidents, 94\% of them, with a precision of 19\%, this corresponds to the high-risk roads and time.
 
@@
 This is not an issue for accident prediction, but it does not help to understand why these roads are particularly dangerous.
-Accident prediction is a very hard machine learning problem because even if the risk is very high for an accident to occur, the driver might manage to avoid it in time and avoid the accident although the conditions were almost exactly the same as when an accident occurs.
+Accident prediction is a very hard machine learning problem because even if the risk is very high for an accident to occur, and even if an accident almost occur, the by reacting quickly, the driver can manage to avoid it although the conditions were almost exactly the same as when an accident occurs.
 We would need information on the state of the drivers and the state of the vehicle to significantly improve performances.
@@
 However, we believe these results can be improved even without this important feature by adding more spatial features.
-We can see that one of the most important variables are the localization of the accident, but the ability to learn a spatial distribution by the random forest algorithm is limited.
+We can see that the most important variables are the features of the road segment.
 Therefore we believe providing more context information on where the road is located will greatly help.
 In another experiment, we have seen the population density and the distance to the closest billboard used as features successfully.
-However, it will be hard to reach good performances because of the nature of the problem without the count of accidents.
+However, even with more features and more data, it will be hard to reach very good performances because of the nature of the problem.
 

### 2019-04-21 paper/paper.tex
Add first part of new related work

@@
 
+need to be extended:
+  how our dataset is bigger because 6 years and precise because road segments
+
 \subsection{Related work}
 Accident prediction has been extensively studied in the last decade.
-The main techniques that have been used at the beginning of the research in road accidents prediction were methods using the Poisson distribution and the negative binomial regression \cite{Chang2005}.
-Then, as machine learning methods become more popular, algorithms like decision trees have been used in several forms like classification and regression decision trees, frequent pattern trees and random forests
-\cite{Chang2005, Theofilatos2017, Abellan2013, Lin2015}, along with other machine learning algorithms like Support Vector Machines \cite{LI20081611}.
+Historically, variation of the Poisson regression such as the negative binomial regression were used to predict the number of accident which occurred on a given road segment.
+The features usually used include feature from the road such as the number of lanes, the average daily traffic, and the curvature, and weather features such as average precipitation and the temperature.
+In 2005, Chang \cite{Change2005} compared the performances of a negative binomial regression with that of an Artificial Neural Network to predict the number of accident during a year on a road segment from a major freeway in Taiwan.
+The dataset used contains data from the years 1997 and 1998, which resulted in 1338 accidents.
+The ANN achieved slightly better result, with an accuracy of 61.4\% on the test dataset.
+On the same dataset, Chang et al.\cite{Chang2005b} also used decision trees for accident prediction in order to get more insight on the important variable for accident prediction.
+It appears that the average daily traffic and the number of days with precipitation are the most useful features.
+The decision tree reached an accuracy of 52.6\% on the test dataset.
+Lin et al.
+\cite{Lin2015} compared Frequent Pattern trees and random forests for feature selection and then k nearest neighbor and bayesian network for real-time accident prediction.
+Using the mean and sometimes the standard deviation of time series corresponding to the weather condition, the visibility, the traffic volume, the traffc speed and the occupancy, their models predict the occurence of an accident at the end of the time series.
+They obtain the best result using the Frequent Pattern trees feature selection and the  (add accuracy)
+We note that the used a small sample of the possible negative examples.
+
+  In addition, some other work aim at predicting the serverity of an accident using various information from the accident in order to understand what causes an accident to be fatal.
+Chong et al.
+\cite{Chong2005} used decision trees, neural network and an hybrid model using a decision tree and a neural network.
+They obtained thes best performances with the hybrid model which reached an accuracy of 90\% for the prediction of fatal injuries.
+They identified that the seatbelt usage, the light conditions and the alcohol usage of the driver are the most important features.
+
+To be continued.
+To cite: \cite{Theofilatos2017, Abellan2013, Lin2015}
+
 Recently, the outstanding progress of deep learning algorithms has resulted in their extensive use in every field, including in Geospatial machine learning (Convolutional Neural Network and Long Short-Term Memory neural network for example) \cite{Yuan2018, QChen2016}.
+
 \section{Methods}
@@
 
-We obtain an area under the precision-recall curve of 54\% and an area under the ROC curve of 89\%.
+We obtain an area under the precision-recall curve of 54\% and an area under the ROC curve of 89\% on the test dataset.
 This results in the following precision and recall values depending on the chosen probability threshold:
@@
 We would need information on the state of the drivers and the state of the vehicle to significantly improve performances.
+
+As opposed to what we would have expected, the Balance Random Forest did not obtained much better performances than the Random Forests with random undersampling.
+We believe this is caused by the fact that negative examples are not that different from each other and that the information they contains is well captured by a single random subsample.
+We can notice, that the Balance Random FOrest does not overfit as much as the Random Forest algorithm.
 

### 2019-04-22 paper/paper.tex
Paper: Rewrote Choice of an evaluation metric: give more explanations

@@
 
-\subsection{Choice of a metric to evaluate our models}
-The chosen metric for the evaluation of the models was the area under the precision-recall curve (area under PR).
-According to Davis and Goadrich\cite{Davis} this measure is the most representative of how well a model has been trained in the case of data imbalance, particularly compared to the ROC curve which is the commonly used measure.
-One can intuitively be convinced because in the ROC curve the false positive rate (FPR) is used instead of the precision.
-The precision refers to the proportion of examples detected as positive which are actually positive, while the FPR refers to the proportion of negative detected as positives.
-With an imbalanced dataset, the FPR will usually stay very low because the number of negative is high.
-The precision, by contrast, will usually be quite low since algorithms tuned for data imbalance will usually include many false positives in the examples they detect in order to achieve a correct recall also called true positive rate.
+\subsection{Choice of an evaluation metric}
+Usually, the accuracy is used to evaluate the performance of classification models, but with an imbalance dataset, this measure does not reflect the real predictive power of the model.
+Indeed, it is usually easy to predict that an example is in the negative class and machine learning models will usually achieve a very good accuracy for the negative class.
+Since most examples belong to the negative class, the overall accuracy will be very good as well.
+However, what we are usually interested in, and this is the case in accident prediction, is to correctly detect the positive examples.
+Therefore, we used the precision and the recall measures.
+The precision is the proportion of examples detected as positive which are actually positive and the recall, also called the true positive rate, is the proportion of positive examples which are detected as positive.
+The false positive rate (FPR) is also usually used in binary classification, it is the proportion of negative examples which are detected as positives.
+With an imbalance dataset, the FPR will usually stay very low because the number of negative is high, that is why we used the precision instead.
+By decreasing the threshold used by the machine learning model to consider an example as positive, we can get a better recall with a worse precision, in order to measure the predictive power of our model independently of the choice of the threshold, we used the area under the precision-recall curve as our main metric.
 

### 2019-04-22 paper/paper.tex
Paper: progress on related work

@@
 
-\subsection{Related work}
-Accident prediction has been extensively studied in the last decade.
-Historically, variation of the Poisson regression such as the negative binomial regression were used to predict the number of accident which occurred on a given road segment.
-The features usually used include feature from the road such as the number of lanes, the average daily traffic, and the curvature, and weather features such as average precipitation and the temperature.
+The rest of this paper is organized as follows: section 2 presents the related work of this paper, the section 3 presents the datasets we used and how we fused them to create positive and negative examples for the road accident prediction, the section 4 presents how we performed the feature engineering, the feature selection and the hyper-parameter tuning, the section 5 present our results and the section 6 discusses them.
+
+\section{Related work}
+Accident prediction has been extensively studied in the last decades.
+Historically, variation of the Poisson regression such as the negative binomial regression were used to predict the number of accident which occurred on a given road segment, but during the last decade machine learning algorithms have been use with success for accident prediction.
+The features usually used include feature from the road such as the number of lanes, the average daily traffic, and the curvature, as well as weather features such as average precipitation and the temperature.
 In 2005, Chang \cite{Change2005} compared the performances of a negative binomial regression with that of an Artificial Neural Network to predict the number of accident during a year on a road segment from a major freeway in Taiwan.
@@
 Lin et al.
-\cite{Lin2015} compared Frequent Pattern trees and random forests for feature selection and then k nearest neighbor and bayesian network for real-time accident prediction.
+\cite{Lin2015} compared the performanced of Frequent Pattern trees with that of random forests for feature selection using then k nearest neighbor and bayesian network for real-time accident prediction.
 Using the mean and sometimes the standard deviation of time series corresponding to the weather condition, the visibility, the traffic volume, the traffc speed and the occupancy, their models predict the occurence of an accident at the end of the time series.
-They obtain the best result using the Frequent Pattern trees feature selection and the  (add accuracy)
-We note that the used a small sample of the possible negative examples.
+They obtain the best result using the Frequent Pattern trees feature selection and achieved an accuracy of 61.7\%.
+We note that they used a small sample of the possible negative examples.
+Theofilatos\cite{Theofilatos2017} also used real-time data on two urban arterials of the city of Athens to study road accident likelihood and severity.
+Random Forests were used for feature selection and a Bayesian logistic regression for accident likelyhood prediction.
+The most important features identified were the coefficient of variation of the flow per lane, the speed and the occupancy, and the standard deviation of the speed and the occupancy.
 
-  In addition, some other work aim at predicting the serverity of an accident using various information from the accident in order to understand what causes an accident to be fatal.
+  In addition, many publications aim at predicting the severity of an accident using various information from the accident in order to understand what causes an accident to be fatal.
 Chong et al.
@@
 They identified that the seatbelt usage, the light conditions and the alcohol usage of the driver are the most important features.
+Abellán et al.
+\cite{Abellan2013} also studied traffic accident severity by looking at the decision rules of a decision tree using a dataset of 1801 highway accidents.
+They found that the the type and cause of the accident the light condition, the sex of the driver and the weather were the most important features.
 
-To be continued.
-To cite: \cite{Theofilatos2017, Abellan2013, Lin2015}
+All of these studies use relatively small datasets using data from only a few years or only a few roads.
+Indeed, it can be hard to collect all the necesssary informations to perform road accident prediction on a larger scale, and dealing with big dataset is more difficult.
+However, more recent studies performed accident prediction at a much larger scale, usually using deep learning models.
+Deep learning models can be trained online so that the whole dataset does not need to stay in memory.
+This makes it easier to deal with big datasets.
 
-Recently, the outstanding progress of deep learning algorithms has resulted in their extensive use in every field, including in Geospatial machine learning (Convolutional Neural Network and Long Short-Term Memory neural network for example) \cite{Yuan2018, QChen2016}.
+Chen et al.
+\cite{QChen2016} used human mobility information coming from mobile phone GPS data and historical accident records to build a model for real-time prediction of traffic accident risk in areas of 500 by 500 meters.
+The risk level of an area is defined as the sum of the severity of accidents that occured in the area during the hour.
+Their model achieves an RMSE of 1.0.
+They compared the performance of their deep learning model with the performances of some of the classical machine learning algorithms: a decision tree, a logistic regression and a Support Vector Machine (SVM), which all got worse performances of respectively 1.41, 1.41 and 1.73.
+We note that they have not tried the random forest algorithm which usually has good prediction performances.
+Najjar et al.
+\cite{Najjar2017}, trained a convolutional neural network using historical accident data and satelite images to predict the risk of accident on an intersection using the satellite image of the intersection.
+Their best model reach an accuracy of 73\%.
+
+To cite: \cite{Ren2017, Yuan2018}.
+
+\subsection{Related work on data imbalance ?}
 

### 2019-04-22 paper/paper.tex
Paper: reorganization

@@
 \section{Introduction}
-\subsection{Context}
 According to the World report on road traffic injury prevention, 2004, more than one million people die every year from car accidents and dozens of millions of people are being injured\cite{Peden2004}.
@@
 \begin{itemize}
-\item a demonstration of how simple and easily available datasets can be assembled to obtain meaningful features for road accident prediction,
-\item a comparison of three different machine learning algorithm dealing with data imbalance in the context of road accident prediction,
+\item a demonstration of how simple and easily available datasets can be fused to obtain meaningful features for road accident prediction,
 \item the implementation of Balanced Random Forest introduced by Chen et al.\cite{Chen2004} in Apache Spark for efficient distributed training.
+\item a comparison of three different machine learning algorithms dealing with data imbalance in the context of road accident prediction,
+\item a real-time road accident prediction model for the island of Montreal with good performances
 \end{itemize}
@@
 
-need to be extended:
-  how our dataset is bigger because 6 years and precise because road segments
+\subsection{Need to be extended, some ideas:}
+
+  As compared to previous studies we have a much bigger dataset of more than a hundred thousands accidents and while other recent studies predict the risk  of an accident in large space area of several square kilometers, we try to predict the occurence of an accident in road segments of a few hundred meters.
 
@@
 
-\section{Methods}
+\section{Datasets fusing}
 \subsection{Choosing a Big Data framework: Apache Spark or Dask}
@@
 For all these reasons we found out that Spark fitted our needs best and we switched to it at an early time of the study.
-
-\subsection{Algorithms selected}
-For this study, we chose to focus on tree-based algorithms because they have proven their efficiency compared to classical statistical methods and allow for easier interpretability than deep learning algorithms.
-In particular, we implemented the Balanced Random Forests (BRF) algorithm in Apache Spark to deal with class imbalance which is a prominent issue in road accident prediction.
-To validate the performance of our version, we run an experiment with a simple synthetic dataset to compare its performances with the standard implementation of Random Forest (RF) in Spark using random under-sampling of the majority class.
-As expected, Balanced Random Forest performed better.
-Balanced Random Forests in one of the two methods proposed by Chen, Liaw, and Breiman\cite{Chen2004} to deal with class imbalance when using Random Forest.
-Balanced Random Forests are like Random Forests, but with a difference during the bootstrapping phase, for each tree of the forest, a random under-sampling of the majority class is performed in order to obtain a balanced sample.
-Intuitively, Balance Random Forest is an adaptation of random undersampling of the majority class making use of the fact that Random Forests are an ensemble method.
-Random undersampling usually performs better than more advanced methods like SMOTE or NearMiss \cite{Branco2016}.
-Chen, Liaw and Breiman\cite{Chen2004} also proposed in the same paper, the Weighted Random Forest (WRF) which consists in giving more weight to the minority class at the building time of each tree and during the measure of the gain in impurity and during the class prediction of each terminal node.
-While neither method is clearly better than the other in terms of prediction power, we chose BRF because it is found to be more computationally efficient (because of the under-sampling).
-Interestingly, Wallace et al.
-\cite{Wallace2011} presents a theoretical analysis of the data imbalance problem and suggest to use methods like Balance Random Forest.
-
-We decided to compare the performance of the Balanced Random Forests algorithm with the performances of one of the very efficient implementations of Gradient Boosted Trees: XGBoost which offers parameters to deal with data imbalance as well.
 
@@
 
+\section{Model development}
+
+\subsection{Algorithms selected}
+For this study, we chose to focus on tree-based algorithms because they have proven their efficiency compared to classical statistical methods and allow for easier interpretability than deep learning algorithms.
+In particular, we implemented the Balanced Random Forests (BRF) algorithm in Apache Spark to deal with class imbalance which is a prominent issue in road accident prediction.
+To validate the performance of our version, we run an experiment with a simple synthetic dataset to compare its performances with the standard implementation of Random Forest (RF) in Spark using random under-sampling of the majority class.
+As expected, Balanced Random Forest performed better.
+Balanced Random Forests in one of the two methods proposed by Chen, Liaw, and Breiman\cite{Chen2004} to deal with class imbalance when using Random Forest.
+Balanced Random Forests are like Random Forests, but with a difference during the bootstrapping phase, for each tree of the forest, a random under-sampling of the majority class is performed in order to obtain a balanced sample.
+Intuitively, Balance Random Forest is an adaptation of random undersampling of the majority class making use of the fact that Random Forests are an ensemble method.
+Random undersampling usually performs better than more advanced methods like SMOTE or NearMiss \cite{Branco2016}.
+Chen, Liaw and Breiman\cite{Chen2004} also proposed in the same paper, the Weighted Random Forest (WRF) which consists in giving more weight to the minority class at the building time of each tree and during the measure of the gain in impurity and during the class prediction of each terminal node.
+While neither method is clearly better than the other in terms of prediction power, we chose BRF because it is found to be more computationally efficient (because of the under-sampling).
+Interestingly, Wallace et al.
+\cite{Wallace2011} presents a theoretical analysis of the data imbalance problem and suggest to use methods like Balance Random Forest.
+
+We decided to compare the performance of the Balanced Random Forests algorithm with the performances of one of the very efficient implementations of Gradient Boosted Trees: XGBoost which offers parameters to deal with data imbalance as well.
+
+\subsection{Implementation of balanced random forest in Apache Spark}
+The Balanced Random Forests algorithm was not implemented in Apache Spark.
+An implementation is available in the python library imbalanced-learn\cite{imbalance} which implements many algorithms to deal with data imbalance using an API inspired by scikit-learn, but the size of our dataset made it impossible for us to use this library.
+Therefore, we implemented Balance Random Forests in Apache Spark.
+
+In the Apache Spark implementation of Random Forests, the bootstrap step is made before starting to grow any tree.
+For each sample, an array contains the number of time it will appear in each tree.
+When doing sampling with replacement, each value of this array is given by a Poisson distribution.
+The parameter of the Poisson distribution corresponds to the subsampling rate hyper-parameter of the Random Forest, which specifies the size of the sample used for training each tree as a fraction of the total size of the dataset.
+Indeed, if for example we want each tree to use a sample of the same size of the whole dataset, the subsampling ratio will be set to 1.0, which is indeed the number of time a given example will appear in a tree on average.
+
+In order to implement Balanced Random Forests, we modified the parameter of the Poisson distribution to use the class weight multiplied by the subsampling ratio, this way a negative sample with for example the weight 0.25 has 4 times less chance to be chosen to appear in a given tree.
+This implementation has the advantage that it did not require a big code change and will be easy to test, but it also has a disadvantage, users probably expect linearly correlated weight to be equivalent, which is not the case in our implementation.
+
+In order to be compatible with other possible use cases, the weights are actually applied per samples and not per class.
+This is a choice made by Apache Spark developers that we respected.
+In order to support sample weights, we create a new Poisson distribution for each sample.
+To make sure the random number generator is not reseeded for each sample, we use the same underlying random number generator for all Poisson distributions, this also helps reduce the cost of creating a new Poisson distribution object.
+Like with other estimators accepting weights, our Balanced Random Forests implementation use weight from a weight column in the samples data frame.
+We adapted the Python wrapper of the random forest classifier to accept and forward weights to the algorithm in Scala.
+Finally, we did a first validation of our implementation using an artificially generated imbalanced dataset and compared the results of both standard random forest and BRF.
+While BRF had better results in terms of area under PR (which was our selected measure) (98.\% against 98.1\%), the standard random forest had better F1 score (99.7\% against 98.8\%) with default threshold.
+This emphasized that we were right in our choice of the area under PR curve as our measure in the case of class imbalance, indeed with a good threshold, the BRF model performs better.
+
 \subsection{Feature Engineering}
@@
 
-\subsection{Implementation of balanced random forest in Apache Spark}
-The Balanced Random Forests algorithm was not implemented in Apache Spark.
-An implementation is available in the python library imbalanced-learn\cite{imbalance} which implements many algorithms to deal with data imbalance using an API inspired by scikit-learn, but the size of our dataset made it impossible for us to use this library.
-Therefore, we implemented Balance Random Forests in Apache Spark.
-
-In the Apache Spark implementation of Random Forests, the bootstrap step is made before starting to grow any tree.
-For each sample, an array contains the number of time it will appear in each tree.
-When doing sampling with replacement, each value of this array is given by a Poisson distribution.
-The parameter of the Poisson distribution corresponds to the subsampling rate hyper-parameter of the Random Forest, which specifies the size of the sample used for training each tree as a fraction of the total size of the dataset.
-Indeed, if for example we want each tree to use a sample of the same size of the whole dataset, the subsampling ratio will be set to 1.0, which is indeed the number of time a given example will appear in a tree on average.
-
-In order to implement Balanced Random Forests, we modified the parameter of the Poisson distribution to use the class weight multiplied by the subsampling ratio, this way a negative sample with for example the weight 0.25 has 4 times less chance to be chosen to appear in a given tree.
-This implementation has the advantage that it did not require a big code change and will be easy to test, but it also has a disadvantage, users probably expect linearly correlated weight to be equivalent, which is not the case in our implementation.
-
-In order to be compatible with other possible use cases, the weights are actually applied per samples and not per class.
-This is a choice made by Apache Spark developers that we respected.
-In order to support sample weights, we create a new Poisson distribution for each sample.
-To make sure the random number generator is not reseeded for each sample, we use the same underlying random number generator for all Poisson distributions, this also helps reduce the cost of creating a new Poisson distribution object.
-Like with other estimators accepting weights, our Balanced Random Forests implementation use weight from a weight column in the samples data frame.
-We adapted the Python wrapper of the random forest classifier to accept and forward weights to the algorithm in Scala.
-Finally, we did a first validation of our implementation using an artificially generated imbalanced dataset and compared the results of both standard random forest and BRF.
-While BRF had better results in terms of area under PR (which was our selected measure) (98.\% against 98.1\%), the standard random forest had better F1 score (99.7\% against 98.8\%) with default threshold.
-This emphasized that we were right in our choice of the area under PR curve as our measure in the case of class imbalance, indeed with a good threshold, the BRF model performs better.
-
-\subsection{Hyper-parameter tunings}
-In order to determine the optimal hyperparameter, we first performed automatic hyperparameter tuning using Apache Spark ML library grid search using cross-validation.
-Because the training on the whole dataset would have been too high, we took a small sample of the dataset.
-Still, We could not test many parameters combination using this method.
-
-Once we got a first tuning result with grid search we continued manually by following a plan, do , check, adjust method, using the measure of the area under the precision-recall curve, the precision and recall with different threshold on the test set and the training set we were able to better understand how the performances of our model could be improved.
-
-Interestingly, despite using many trees, our Random Forest algorithms tended to overfit very quickly as soon as the maximum depth parameter goes above 18.
-We eventually used only 100 trees, because adding more trees did not increase performances.
-We have not tried more than 200 trees, maybe many more trees would have been necessary to increase the maximum depth without overfitting, but then the training time would not be reasonable.
+The following table present the statistical characteristic of our features in the traning dataset
 
@@
 \end{figure}
+
+\subsection{Hyper-parameter tunings}
+In order to determine the optimal hyperparameter, we first performed automatic hyperparameter tuning using Apache Spark ML library grid search using cross-validation.
+Because the training on the whole dataset would have been too high, we took a small sample of the dataset.
+Still, We could not test many parameters combination using this method.
+
+Once we got a first tuning result with grid search we continued manually by following a plan, do , check, adjust method, using the measure of the area under the precision-recall curve, the precision and recall with different threshold on the test set and the training set we were able to better understand how the performances of our model could be improved.
+
+Interestingly, despite using many trees, our Random Forest algorithms tended to overfit very quickly as soon as the maximum depth parameter goes above 18.
+We eventually used only 100 trees, because adding more trees did not increase performances.
+We have not tried more than 200 trees, maybe many more trees would have been necessary to increase the maximum depth without overfitting, but then the training time would not be reasonable.
 
@@
 
+Add discussion about why accuracy does not make sense to evaluate performance, and also practical example of what our prediction results means.
+
 \subsection{Future work}
@@
 These datasets might not be as easily available for over areas, but for Montreal, it could improve performances.
-Also, the solar elevation and solar azimuth could easily be added without the need for another dataset.
-These features have been used successfully in road accident prediction in previous experiments.
+This accident prediction model can be used direcly to know on which road segments accidents are the most likely to happen in a given hour in order to take measure to reduce risk.
+This would require to put this model into production.
+The historical climate dataset, would have to be replaced by a weather forecast dataset.
+The most important feature is the number of accident which happened during the previous year, while this feature help a lot to reach useful prediction performances, it does not help in understanding the characteristics of a road segment which makes it dangerous.
+A human analysis of these particularly risky road segment could detect patterns that could help to take measure to reduce the number of accident in Montreal.
+This can also allow to improve our current accident prediction model, if the detected patterns can be used by fusing other datasets.
+
+Add these features:
+Also, the solar elevation and solar azimuth could easily be added without the need for another dataset. These features have been used successfully in road accident prediction in previous experiments.
 
@@
 
-We acknowledge Compute Canada for giving us access to the cluster.
+The authors would like to acknowledge Compute Canada for providing access to the computation cluster.
 

### 2019-04-22 random_forest.py
Implement more efficient version of compute_precision_recall_graph function

@@ -5,12 +5,13 @@ from pyspark.ml.tuning import TrainValidationSplit, \
                               ParamGridBuilder, \
                               CrossValidator
 from pyspark.ml import Pipeline
+from pyspark.sql import Window
 from pyspark.sql.types import DoubleType
-from pyspark.sql.functions import udf, col
+from pyspark.sql.functions import udf, col, floor, sum
 import pandas as pd
 import numpy as np
 from preprocess import features_col
-from random_undersampler import RandomUnderSampler 
+from random_undersampler import RandomUnderSampler
 from class_weighter import ClassWeighter
 
 
@@ -35,10 +36,10 @@ def random_forest_tuning(train_samples):
                                       rawPredictionCol="rawPrediction",
                                       metricName="areaUnderPR")
     tvs = TrainValidationSplit(estimator=pipeline,
-                         estimatorParamMaps=paramGrid,
-                         evaluator=pr_evaluator,
-                         trainRatio=0.8,
-                         collectSubModels=True)
+                               estimatorParamMaps=paramGrid,
+                               evaluator=pr_evaluator,
+                               trainRatio=0.8,
+                               collectSubModels=True)
 
     model = tvs.fit(train_samples)
 
@@ -77,7 +78,6 @@ def balanced_random_forest_tuning(train_samples):
     model = tvs.fit(train_samples)
 
     return model
-
 
 
 def compute_precision_recall(predictions, threshold):
@@ -119,7 +119,7 @@ def compute_precision_recall(predictions, threshold):
     return (precision, recall)
 
 
-def compute_precision_recall_graph(predictions, n_points):
+def compute_precision_recall_graph_slow(predictions, n_points):
     def gen_row(threshold):
         result = compute_precision_recall(predictions, threshold)
         return (threshold, result[0], result[1])
@@ -129,6 +129,43 @@ def compute_precision_recall_graph(predictions, n_points):
                          columns=['Threshold', 'Precision', 'Recall'])
 
     return graph
+
+
+def compute_precision_recall_graph(predictions, n_points):
+    inf_cumulative_window = \
+        (Window
+         .partitionBy('label')
+         .orderBy('id_bucket')
+         .rowsBetween(Window.unboundedPreceding, Window.currentRow))
+    sup_cumulative_window = \
+        (Window
+         .partitionBy('label')
+         .orderBy('id_bucket')
+         .rowsBetween(1, Window.unboundedFollowing))
+
+    return \
+        (predictions
+         .select('label',
[…]

### 2019-04-22 accidents_montreal.py
Fix code style

@@ -13,6 +13,7 @@ from utils import raise_parquet_not_del_error
 from shutil import rmtree
 from workdir import workdir
 
+
 def get_accident_df(spark, use_cache=True):
     fetch_accidents_montreal()
     return read_accidents_montreal_df(spark, use_cache)

### 2019-04-22 class_weighter.py

@@ -9,6 +9,7 @@ class HasClassWeight(Params):
                         "classWeight",
                         "Array containing the weight to give to each class",
                         typeConverter=TypeConverters.toListFloat)
+
     def __init__(self):
         super(HasClassWeight, self).__init__()
 
@@ -36,4 +37,3 @@ class ClassWeighter(Transformer, HasWeightCol, HasLabelCol, HasClassWeight):
                                   when(col(self.getLabelCol()) == 0.0,
                                        lit(class_weight[0]))
                                   .otherwise(lit(class_weight[1])))
-

### 2019-04-22 main_brf.py

@@ -13,8 +13,8 @@ from workdir import workdir
 
 spark = init_spark()
 
-i=1
-sampleFraction=0.01
+i = 1
+sampleFraction = 0.01
 
 neg_samples = get_negative_samples(spark).sample(sampleFraction)
 pos_samples = get_positive_samples(spark).sample(sampleFraction)

### 2019-04-22 main_negative_samples.py

@@ -3,7 +3,8 @@ from preprocess import get_negative_samples, get_positive_samples
 from utils import init_spark
 from workdir import workdir
 spark = init_spark()
-neg_samples = get_negative_samples(spark,
-                                   save_to=workdir + 'data/negative-sample-new.parquet',
-                                   sample_ratio=1e-3)
+neg_samples = \
+    get_negative_samples(spark,
+                         save_to=workdir + 'data/negative-sample-new.parquet',
+                         sample_ratio=1e-3)
 print(neg_samples.count())

### 2019-04-22 main_xgboost.py

@@ -45,10 +45,12 @@ def grid_search(spark):
 
     pipeline = Pipeline().setStages([xgboost])
     model = pipeline.fit(trainDF)
-    paramGrid = (ParamGridBuilder().addGrid(xgboost.max_depth, [x for x in range(3, 20, 6)])
-                                   .addGrid(xgboost.eta, [x for x in np.linspace(0.2, 0.6, 4)])
-                                   .addGrid(xgboost.scale_pos_weight, [x for x in np.linspace(0.03, 1.0, 3)])
-                                   .build())
+    paramGrid = (ParamGridBuilder()
+                 .addGrid(xgboost.max_depth, [x for x in range(3, 20, 6)])
+                 .addGrid(xgboost.eta, [x for x in np.linspace(0.2, 0.6, 4)])
+                 .addGrid(xgboost.scale_pos_weight,
+                          [x for x in np.linspace(0.03, 1.0, 3)])
+                 .build())
     evaluator = BinaryClassificationEvaluator(labelCol="label",
                                               rawPredictionCol="probabilities",
                                               metricName="areaUnderPR")
@@ -75,7 +77,8 @@ def grid_search(spark):
             file.write(f"Area under PR: {result}\n")
 
     prediction = bestModel.transform(testDF)
-    prediction = prediction.withColumn("rawPrediction", prediction['probabilities'])
+    prediction = prediction.withColumn("rawPrediction",
+                                       prediction['probabilities'])
     area_under_PR, f1_score = evaluate_binary_classifier(prediction)
 
     with open(workdir + 'data/xgboost_tuning_perf_1.txt', 'w') as file:
@@ -86,5 +89,5 @@ def grid_search(spark):
 
 spark = init_spark()
 spark.sparkContext.addPyFile(workdir + "data/sparkxgb.zip")
-from sparkxgb import XGBoostEstimator
+from sparkxgb import XGBoostEstimator  # noqa
 grid_search(spark)

### 2019-04-22 preprocess.py

@@ -429,8 +429,10 @@ def get_dataset_df(spark, pos_samples, neg_samples):
 
     pos_samples = pos_samples.select(*neg_samples.columns)
     df = pos_samples.union(neg_samples)
-    street_lvl_index, df = index_street_categories(spark, pos_samples, df, 'street_level')
-    street_typ_index, df = index_street_categories(spark, pos_samples, df, 'street_type')
+    street_lvl_index, df = \
+        index_street_categories(spark, pos_samples, df, 'street_level')
+    street_typ_index, df = \
+        index_street_categories(spark, pos_samples, df, 'street_type')
 
     assembler = VectorAssembler(outputCol="features",
                                 inputCols=features_col,
@@ -441,18 +443,20 @@ def get_dataset_df(spark, pos_samples, neg_samples):
                   'street_id',
                   'date', 'hour', 'features', 'label'))
 
-    return street_lvl_index, street_typ_index, df.withColumn('id', monotonically_increasing_id())
+    return (street_lvl_index,
+            street_typ_index,
+            df.withColumn('id', monotonically_increasing_id()))
 
 
 def index_street_categories(spark, pos_samples, df, col_name):
     road = get_road_features_df(spark)
     street_cat_distrib = (road
-                           .select(col_name)
-                           .na.fill("unknown")
-                           .groupBy(col_name).count()
-                           .withColumn('p', col('count') / lit(road.count()))
-                           .orderBy(col('p').desc())
-                           .drop('count'))
+                          .select(col_name)
+                          .na.fill("unknown")
+                          .groupBy(col_name).count()
+                          .withColumn('p', col('count') / lit(road.count()))
+                          .orderBy(col('p').desc())
+                          .drop('count'))
     street_cat_indexes = \
         (pos_samples
          .select(col_name)
@@ -465,6 +469,7 @@ def index_street_categories(spark, pos_samples, df, col_name):
          .select(col_name, (col('p_pos') - col('p')).alias('risk'))
          .withColumn(col_name + '_indexed',
                      row_number().over(Window.orderBy(col('risk').desc())))
-        . drop('risk'))
+         .drop('risk'))
 
-    return street_cat_indexes, df.join(street_cat_indexes, col_name).drop(col_name)
+    return (street_cat_indexes,
+            df.join(street_cat_indexes, col_name).drop(col_name))

### 2019-04-22 random_undersampler.py

@@ -27,12 +27,16 @@ class HasTargetImbalanceRatio(Params):
         """
         return self._set(targetImbalanceRatio=value)
 
+
 class HasIndexCol(Params):
     """
     Mixin for param indexCol: index column name.
     """
 
-    indexCol = Param(Params._dummy(), "indexCol", "index column name.", typeConverter=TypeConverters.toString)
+    indexCol = Param(Params._dummy(),
+                     "indexCol",
+                     "index column name.",
+                     typeConverter=TypeConverters.toString)
 
     def __init__(self):
         super(HasIndexCol, self).__init__()
@@ -67,6 +71,7 @@ class RandomUnderSampler(Estimator, HasTargetImbalanceRatio, HasSeed,
 
         return (RandomUnderSamplerModel(indexes_to_remove, self.getSeed())
                 .setTargetImbalanceRatio(self.getTargetImbalanceRatio()))
+
 
 class RandomUnderSamplerModel(Transformer, HasTargetImbalanceRatio):
     def __init__(self, indexesToRemove, seed):

### 2019-04-22 road_network.py

@@ -18,7 +18,6 @@ from workdir import workdir
 from road_network_nids import unknow_file_included_nids
 
 
-
 def get_road_df(spark, use_cache=True):
     fetch_road_network()
     return extract_road_segments_df(spark, use_cache)
@@ -84,7 +83,7 @@ def fetch_road_network():
         return
     print('Fetching road network...')
     url = 'http://ftp.maps.canada.ca/pub/nrcan_rncan'\
-                '/vector/geobase_nrn_rrn/qc/kml_en/'
+          '/vector/geobase_nrn_rrn/qc/kml_en/'
     if not os.path.isdir(workdir + 'data/road-network'):
         os.mkdir(workdir + 'data/road-network')
     files = \
@@ -140,10 +139,13 @@ def get_kml_content(soup):
             coordinates_list = (placemark.MultiGeometry.LineString.coordinates
                                 .text.split(' '))
             description = placemark.find('description').text
-            nid = re.search('<th>nid</th>\n<td>([a-f0-9]+)</td>', description).group(1)
-            is_unknown = re.search('<th>left_OfficialPlaceName</th>\n<td>Unknown</td>', description) is not None
+            nid = (re.search('<th>nid</th>\n<td>([a-f0-9]+)</td>', description)
+                   .group(1))
+            is_unknown = \
+                re.search('<th>left_OfficialPlaceName</th>\n<td>Unknown</td>',
+                          description) is not None
             if is_unknown and (nid not in unknow_file_included_nids):
-                    continue
+                continue
 
             for coord in coordinates_list:
                 coords = coord.split(',')
@@ -206,11 +208,13 @@ def extract_road_segments_df(spark, use_cache=True):
 
     # Some specific road segments have the same nid
     w = Window.partitionBy('nid').orderBy('center_lat')
-    street_ids = (road_seg_df
-                  .select('nid', 'center_lat', 'center_long')
-                  .distinct()
-                  .select('center_lat', 'center_long',
-                          concat('nid', row_number().over(w)).alias('street_id')))
+    street_ids = \
+        (road_seg_df
+         .select('nid', 'center_lat', 'center_long')
+         .distinct()
+         .select('center_lat',
+                 'center_long',
+                 concat('nid', row_number().over(w)).alias('street_id')))
 
     road_seg_df = (road_seg_df
                    .join(street_ids, ['center_lat', 'center_long'])

### 2019-04-22 road_network_nids.py

@@ -1,1774 +1,1775 @@
-# List obtained by graphically selection on google earth the road inside the island of Montreal
+# List obtained by graphically selection on google earth the road inside the
+# island of Montreal
 unknow_file_included_nids = ['73671c5348054b448b94358ac26ff548',
-'0de4c13b13334812ba1803a4d2ef1cf8',
-'72d5c5f727614966b8a3733c0a2c032d',
-'c34dadb8dcf24143b3d14eab75dae974',
-'72beb531ce7f476fa383616a7ed7276b',
-'3f26f14865b244a9acf31f812dc83aa0',
-'b26c24b5349442049a74742dd4c9a3f9',
-'6b8c930b1f53480a9eca6365c3c4a7e7',
-'faf1cb7e937d44639440e1eefff5adea',
-'8ec66e8fa96643618cefe8ed8735eef6',
-'e3fa30f1cd8641beab219f00a32d391d',
-'15a3d60020824b2285030b7fa77f0a6c',
-'3c84b65d022a401190a0de49858a5e22',
-'3408fd326abe43af86a61250b1111efa',
-'40086095074e446b96a36865de7a8679',
-'e6c2a08cf30c4566afdfd2832080d5d3',
-'ec0d0ad9a5334d4ca5893638969863f1',
-'d6695b4cc4cf46189eac7561d1fa72a2',
-'b5b3fe8dd24348ad99fcdd0282efc7ab',
-'faf36224560f4f82b18aa08a7b889c8c',
-'4bbcd933bcaf42a3a95b1a69efefb8b4',
-'a67f895a006c425289a84a0225fd8b96',
-'fc0225d59e43435488ea1f814735b6ad',
-'30cd4380cc454645b1a8cb85c0a240b9',
-'d8d16706c88a485884b5284324167076',
-'73d4bf983e8647cc880441e41ea3bc8c',
-'aa0294bb10734b4cb1b91392a7a5402c',
-'0787e804523248bfa73ea28e76b5be96',
-'500f55d39dfe4aeb8465d338974dd78c',
-'e64ba475514d4f48ab8ce2ea044cece4',
-'9721d22f9b8f4d0188bcd3220aa05fe7',
-'51ec453db2104bad933c8cf893a12fba',
[…]

### 2019-04-22 utils.py

@@ -20,18 +20,18 @@ def init_spark():
     sess = (SparkSession
             .builder
             .appName("Accident prediction")
-            #.master("local[6]")
-            #.config("spark.driver.memory", "4g")
+            # .master("local[6]")
+            # .config("spark.driver.memory", "4g")
             .config("spark.rdd.compress", "True")
             .config("spark.serializer",
                     "org.apache.spark.serializer.KryoSerializer")
             # In local mode there is only one executor: the driver
             # .config("spark.executor.memory", "1536m")
-            #.config("spark.driver.memory", "3g")
+            # .config("spark.driver.memory", "3g")
             # Prevent time out errors see: https://github.com/rjagerman/mammoth
             # /wiki/ExecutorLost-Failure:-Heartbeat-timeouts
-            #.config("spark.cleaner.periodicGC.interval", "5min")
-            #.config("spark.network.timeout", "300s")
+            # .config("spark.cleaner.periodicGC.interval", "5min")
+            # .config("spark.network.timeout", "300s")
             .config("spark.driver.extraClassPath",
                     f"{workdir}data/xgboost4j-spark-0.72.jar"
                     + f":{workdir}data/xgboost4j-0.72.jar")
@@ -42,6 +42,7 @@ def init_spark():
     for param in sess.sparkContext.getConf().getAll():
         print(f"\t{param[0]}: {param[1]}")
     return sess
+
 
 def get_with_retry(
     url,

### 2019-04-22 weather.py

@@ -108,8 +108,8 @@ def get_station_weather_month(station_id, year, month):
     df['risky_weather'] = np.nan
     df.loc[df['Temp (°C)'].notna(), 'risky_weather'] = \
         df['Weather'].astype(str).str.contains('snow|ice|freezing',
-                                   regex=True,
-                                   flags=re.IGNORECASE)
+                                               regex=True,
+                                               flags=re.IGNORECASE)
 
     return [(
         int(r['Day']),
@@ -189,23 +189,22 @@ def get_weather_station_weather_df(spark, stations_id):
         def value(i):
             return lag(c, -i).over(window)
 
-        values = [coalesce(value(i) * w, lit(0)) 
-                for i, w in zip(offsets, weights)]
+        values = [coalesce(value(i) * w, lit(0))
+                  for i, w in zip(offsets, weights)]
 
         return sum(values, lit(0))
 
     window = (Window
-            .partitionBy('station_id')
-            .orderBy('date'))
+              .partitionBy('station_id')
+              .orderBy('date'))
     offsets = range(-23, 1)
-    l = 0.5
-    weights = [exp(l*t) for t in offsets]
+    weights = [exp(0.5*t) for t in offsets]
     weights = [w/sum(weights) for w in weights]
     df = df.withColumn('risky_weather',
-                            weighted_average('risky_weather',
-                                            window,
-                                            offsets,
-                                            weights))
+                       weighted_average('risky_weather',
+                                        window,
+                                        offsets,
+                                        weights))
 
     df.write.parquet(cache_file)
 
@@ -228,7 +227,7 @@ def get_weather_station_coords(station_id):
     lat = coords_div.div.div.find_all('div', recursive=False)[1].text
     long = (coords_div.find_all('div', recursive=False)[1].div
             .find_all('div', recursive=False)[1].text)
-    return DMS_to_degree(lat), DMS_to_degree(long)
+    return DMS_to_degree(lat), -DMS_to_degree(long)
 
 
 def get_weather_station_coords_df(spark, stations_id):

### 2019-04-22 preprocess.py
Implement train-test split, accident count features, and street_type/level indexing in get_dataset_df

@@ -428,48 +428,74 @@ def get_dataset_df(spark, pos_samples, neg_samples):
     neg_samples = neg_samples.withColumn('label', lit(0.0))
 
     pos_samples = pos_samples.select(*neg_samples.columns)
-    df = pos_samples.union(neg_samples)
-    street_lvl_index, df = \
-        index_street_categories(spark, pos_samples, df, 'street_level')
-    street_typ_index, df = \
-        index_street_categories(spark, pos_samples, df, 'street_type')
 
-    assembler = VectorAssembler(outputCol="features",
-                                inputCols=features_col,
-                                handleInvalid='keep'
-                                )
-    df = (assembler.transform(df)
-          .select('sample_id',
-                  'street_id',
-                  'date', 'hour', 'features', 'label'))
+    train_pos, train_neg, test_pos, test_neg = \
+        train_test_split(pos_samples, neg_samples)
 
-    return (street_lvl_index,
-            street_typ_index,
-            df.withColumn('id', monotonically_increasing_id()))
+    roads = get_road_features_df(spark)
+    accident_count = get_accident_count(spark, train_pos, roads)
+    street_level_index = \
+        get_street_categories_index(spark, 'street_level', train_pos, roads)
+    street_type_index = \
+        get_street_categories_index(spark, 'street_type', train_pos, roads)
+
+    def prepare_dataset(va, pos_samples, neg_samples):
+        dataset = pos_samples.union(neg_samples)
+        dataset = (dataset
+                   .join(accident_count, 'street_id')
+                   .join(street_level_index, 'street_level')
+                   .join(street_type_index, 'street_type')
+                   .drop('street_level', 'street_type'))
+        return (va.transform(dataset)
+                .select('sample_id', 'street_id', 'date',
+                        'hour', 'features', 'label'))
+
+    va = VectorAssembler(outputCol="features",
+                         inputCols=features_col,
+                         handleInvalid='keep')
+
+    train_set = prepare_dataset(va, train_pos, train_neg)
+    test_set = prepare_dataset(va, test_pos, test_neg)
+
+    return train_set, test_set
 
 
-def index_street_categories(spark, pos_samples, df, col_name):
-    road = get_road_features_df(spark)
-    street_cat_distrib = (road
+def train_test_split(pos_samples, neg_samples):
+    split_date = datetime.datetime.fromisoformat('2017-01-01')
+    train_pos = (pos_samples.filter(col('date') < split_date))
[…]

### 2019-04-22 main.py
Move mains to dedicated directory

@@ -1,9 +1,0 @@
-#!/usr/bin/env python3
-from road_network import get_road_df
-from preprocess import get_positive_samples, get_negative_samples
-from utils import init_spark
-
-spark = init_spark()
-
-print('testing positives generation...')
-positive_samples = get_positive_samples(spark, limit=400)

### 2019-04-22 mains/main_brf.py

@@ -1,4 +1,5 @@
 #!/usr/bin/env python3
+import accident_prediction_montreal
 from datetime import datetime
 from pyspark.sql.functions import col
 from random_forest import balanced_random_forest_tuning, \

### 2019-04-22 mains/main_gen_match.py

@@ -1,4 +1,5 @@
 #!/usr/bin/env python
+import accident_prediction_montreal
 import os
 from functools import reduce
 import datetime

### 2019-04-22 mains/main_make_weather_cache.py

@@ -1,5 +1,5 @@
 #!/usr/bin/env python
-
+import accident_prediction_montreal
 from accidents_montreal import get_accident_df
 from weather import get_weather_df
 from utils import init_spark

### 2019-04-22 mains/main_negative_samples.py

@@ -1,4 +1,5 @@
 #!/usr/bin/env python
+import accident_prediction_montreal
 from preprocess import get_negative_samples, get_positive_samples
 from utils import init_spark
 from workdir import workdir

### 2019-04-22 mains/main_positive_samples.py

@@ -1,4 +1,5 @@
 #!/usr/bin/env python
+import accident_prediction_montreal
 from preprocess import get_positive_samples
 from utils import init_spark
 spark = init_spark()

### 2019-04-22 mains/main_random_forest.py

@@ -1,4 +1,5 @@
 #!/usr/bin/env python3
+import accident_prediction_montreal
 from datetime import datetime
 from pyspark.sql.functions import col
 from random_forest import random_forest_tuning, \

### 2019-04-22 mains/main_weather_stations_id.py

@@ -1,5 +1,5 @@
 #!/usr/bin/env python
-
+import accident_prediction_montreal
 from accidents_montreal import get_accident_df
 from weather import get_useful_stations_id_df
 from utils import init_spark

### 2019-04-22 mains/main_xgboost.py

@@ -2,6 +2,7 @@ import numpy as np
 import shutil
 import os
 
+import accident_prediction_montreal
 from evaluate import *
 from random_forest import *
 from utils import init_spark

### 2019-04-22 paper/paper.tex
Paper: Add newline between each sentence

@@
 Road accidents are an important issue of our modern societies, responsible for millions of deaths and injuries.
-In this paper we show how one can leverage open datasets from a city like Montreal, Canada, to create some accident prediction models, using state-of-the-art big data analytics methods in Python.
+In this paper we show how one can leverage open datasets from a city like Montreal, Canada, to create some accident prediction models, using state-of-the-art big data analytics methods.
 Such models could then be used in the context of road accident prevention, but also to identify key factors that can lead to a road accident, and eventually, help to elaborate new policies.
@@
   In addition, many publications aim at predicting the severity of an accident using various information from the accident in order to understand what causes an accident to be fatal.
-Chong et al.
-\cite{Chong2005} used decision trees, neural network and an hybrid model using a decision tree and a neural network.
+Chong et al.\cite{Chong2005} used decision trees, neural network and an hybrid model using a decision tree and a neural network.
 They obtained thes best performances with the hybrid model which reached an accuracy of 90\% for the prediction of fatal injuries.
 They identified that the seatbelt usage, the light conditions and the alcohol usage of the driver are the most important features.
-Abellán et al.
-\cite{Abellan2013} also studied traffic accident severity by looking at the decision rules of a decision tree using a dataset of 1801 highway accidents.
+Abellán et al. \cite{Abellan2013} also studied traffic accident severity by looking at the decision rules of a decision tree using a dataset of 1801 highway accidents.
 They found that the the type and cause of the accident the light condition, the sex of the driver and the weather were the most important features.
@@
 This format is based on the Extensible Markup Language (XML), which make it easier to read using existing implementation of XML parsers.
+
 \item Data Historical Climate Data: This dataset provided by the government of Canada contains hourly weather information measured at different weather stations across Canada.
@@
 While neither method is clearly better than the other in terms of prediction power, we chose BRF because it is found to be more computationally efficient (because of the under-sampling).
-Interestingly, Wallace et al.
-\cite{Wallace2011} presents a theoretical analysis of the data imbalance problem and suggest to use methods like Balance Random Forest.
-
+Interestingly, Wallace et al. \cite{Wallace2011} presents a theoretical analysis of the data imbalance problem and suggest to use methods like Balance Random Forest.
 We decided to compare the performance of the Balanced Random Forests algorithm with the performances of one of the very efficient implementations of Gradient Boosted Trees: XGBoost which offers parameters to deal with data imbalance as well.
@@
 Add these features:
-Also, the solar elevation and solar azimuth could easily be added without the need for another dataset. These features have been used successfully in road accident prediction in previous experiments.
+Also, the solar elevation and solar azimuth could easily be added without the need for another dataset.
+These features have been used successfully in road accident prediction in previous experiments.
 

### 2019-04-22 random_forest.py
Fix compute_precision_recall_graph

@@ -143,10 +143,19 @@ def compute_precision_recall_graph(predictions, n_points):
          .orderBy('id_bucket')
          .rowsBetween(1, Window.unboundedFollowing))
 
+    def prob_positive(v):
+        try:
+            return float(v[1])
+        except ValueError:
+            return None
+
+    prob_positive = udf(prob_positive, DoubleType())
+
     return \
         (predictions
          .select('label',
-                 floor(col('probability') * n_points).alias('id_bucket'))
+                 floor(prob_positive('probability') * n_points)
+                 .alias('id_bucket'))
          .groupBy('label', 'id_bucket').count()
          .withColumn('count_negatives',
                      sum('count').over(inf_cumulative_window))
@@ -165,7 +174,9 @@ def compute_precision_recall_graph(predictions, n_points):
                  .alias('Precision'),
                  (col('true_positive')
                  / (col('true_positive') + col('false_negative')))
-                 .alias('Recall'))).toPandas()
+                 .alias('Recall'))
+         .orderBy('Threshold')
+         .toPandas())
 
 
 def get_feature_importances(model):

### 2019-04-22 preprocess.py
Add the feature solar elevation in solar_features.py

@@ -19,6 +19,7 @@ from road_network import distance_intermediate_formula,\
                          get_road_features_df,\
                          get_road_df
 from weather import get_weather_df
+from solar_features import add_solar_features
 from workdir import workdir
 
 
@@ -272,6 +273,8 @@ def get_negative_samples(spark, use_cache=True, save_to=None, road_limit=None,
     negative_samples = \
         negative_samples.join(negative_sample_weather, 'sample_id')
     negative_samples = add_date_features(negative_samples)
+    negative_samples = add_solar_features(negative_samples)
+
     negative_samples = negative_samples.persist()
 
     if use_cache:
@@ -316,6 +319,7 @@ def get_positive_samples(spark, road_df=None, weather_df=None,
                         .withColumnRenamed('accident_id', 'sample_id'))
 
     positive_samples = add_date_features(positive_samples)
+    positive_samples = add_solar_features(positive_samples)
     positive_samples = positive_samples.persist()
 
     if use_cache:
@@ -403,7 +407,8 @@ features_col = ['hour_cos',
                 'risky_weather',
                 'dayofyear_cos',
                 'dayofyear_sin',
-                # 'dayofweek',
+                'dayofweek',
+                'solar_elevation',
                 # 'dayofmonth_cos',
                 # 'dayofmonth_sin',
                 ]

### 2019-04-22 solar_features.py

@@ -1,0 +1,52 @@
+from datetime import datetime
+from math import pi
+from pyspark.sql.functions import col, cos, sin, asin, when, dayofyear, \
+                                  radians, degrees
+
+
+def add_solar_features(df):
+    return \
+        (df
+         .withColumn('declination_angle',
+                     radians(-23.45
+                             * cos(((2 * pi)/365) * (dayofyear('date') + 10))))
+         .withColumn('diff_local_time_UTC', timezone_from_date('date'))
+         .withColumn('d', (2 * pi * dayofyear('date')) / 365)
+         .withColumn('equation_of_time',
+                     -7.655 * sin(col('d'))
+                     + 9.873 * sin(2 * col('d') + 3.588))
+         .drop('d')
+         .withColumn('time_correction',
+                     4 * (col('loc_long') - (15 * col('diff_local_time_UTC')))
+                     + col('equation_of_time'))
+         .withColumn('local_solar_hour',
+                     col('hour') + 0.5 + col('time_correction') / 60)
+         .withColumn('hour_angle', 0.2618 * (col('local_solar_hour') - 12))
+         .drop('diff_local_time_UTC', 'equation_of_time', 'time_correction',
+               'local_solar_hour')
+         .withColumn('solar_elevation',
+                     degrees(asin(sin('declination_angle')
+                                  * sin(radians('loc_lat'))
+                                  + cos('declination_angle')
+                                  * cos(radians('loc_lat'))
+                                  * cos('hour_angle'))))
+         .drop('declination_angle', 'hour_angle'))
+
+
+def timezone_from_date(date):
+    todate = datetime.fromisoformat
+    return when(((col(date) > todate('2012-03-11 02:00:00'))
+                & (col(date) < todate('2012-11-04 02:00:00')))
+                | ((col(date) > todate('2013-03-10 02:00:00'))
+                & (col(date) < todate('2013-11-03 02:00:00')))
+                | ((col(date) > todate('2014-03-09 02:00:00'))
+                & (col(date) < todate('2014-11-02 02:00:00')))
+                | ((col(date) > todate('2015-03-08 02:00:00'))
+                & (col(date) < todate('2015-11-01 02:00:00')))
+                | ((col(date) > todate('2016-03-13 02:00:00'))
+                & (col(date) < todate('2016-11-06 02:00:00')))
+                | ((col(date) > todate('2017-03-12 02:00:00'))
[…]

### 2019-04-23 evaluate.py
Add more evaluation metrics

@@ -13,6 +13,20 @@ def evaluate_binary_classifier(predictions):
                                           labelCol='label',
                                           metricName='f1')
     f1_score = f1_evaluator.evaluate(predictions)
-    print(f"Area Under PR = {area_under_PR}\nF1 score = {f1_score}")
+    ROC_evaluator = \
+        BinaryClassificationEvaluator(labelCol="label",
+                                      rawPredictionCol="rawPrediction",
+                                      metricName="areaUnderROC")
+    area_under_ROC = ROC_evaluator.evaluate(predictions)
+    acc_evaluator = \
+    MulticlassClassificationEvaluator(predictionCol='prediction',
+                                      labelCol='label',
+                                      metricName='accuracy')
+    acc_score = acc_evaluator.evaluate(predictions)
 
-    return (area_under_PR, f1_score)
+    print(f"Area Under PR = {area_under_PR}")
+    print(f"F1 score = {f1_score}")
+    print(f"Area Under ROC = {area_under_ROC}")
+    print(f"Accuracy = {acc_score}")
+
+    return (area_under_PR, f1_score, area_under_ROC, acc_score)

### 2019-04-23 random_forest.py
Fix get_feature_importances

@@ -181,9 +181,7 @@ def compute_precision_recall_graph(predictions, n_points):
 
 def get_feature_importances(model):
     feature_importances = pd.DataFrame(model.featureImportances.toArray())
-    dayofweek_features = [f'{features_col[-1]}_{i}' for i in range(1, 8)]
-    feature_names = features_col[:-1] + dayofweek_features
-    feature_importances.index = feature_names
+    feature_importances.index = features_col
     feature_importances.columns = ["Feature importances"]
     feature_importances.sort_values(['Feature importances'], ascending=False)
     return feature_importances

### 2019-04-23 mains/main_train_brf.py
Implement BRF training script

@@ -1,0 +1,85 @@
+#!/usr/bin/env python
+from os import isdir, mkdir
+from preprocess import get_negative_samples, get_positive_samples
+from utils import init_spark
+from preprocess import get_dataset_df
+from pyspark.ml.classification import RandomForestClassifier
+from pyspark.ml.evaluation import BinaryClassificationEvaluator
+from pyspark.ml.evaluation import MulticlassClassificationEvaluator
+from pyspark.ml.tuning import ParamGridBuilder, TrainValidationSplit, CrossValidator
+from pyspark.ml import Pipeline
+from random_undersampler import RandomUnderSampler
+from class_weighter import ClassWeighter
+from workdir import workdir
+from random_forest import compute_precision_recall_graph
+from random_forest import get_feature_importances
+from evaluate import evaluate_binary_classifier
+
+# Prepare result directory
+i = 1
+while isdir(workdir + f'data/brf_{i}'):
+    i += 1
+mkdir(workdir + f'data/brf_{i}')
+
+spark = init_spark()
+neg_samples = get_negative_samples(spark).sample(0.5)
+pos_samples = get_positive_samples(spark)
+
+imbalance_ratio = (neg_samples.count()/pos_samples.count())
+
+train_set, test_set = get_dataset_df(spark, pos_samples, neg_samples)
+train_set, test_set = train_set.persist(), test_set.persist()
+
+brf = RandomForestClassifier(labelCol="label",
+                             featuresCol="features",
+                             cacheNodeIds=True,
+                             maxDepth=17,
+                             impurity='entropy',
+                             featureSubsetStrategy='sqrt',
+                             weightCol='weight',
+                             minInstancesPerNode=10,
+                             numTrees=100,
+                             subsamplingRate=1.0,
+                             maxMemoryInMB=768)
+cw = ClassWeighter().setClassWeight([1/imbalance_ratio, 1.0])
+pipeline = Pipeline().setStages([cw, brf])
+model = pipeline.fit(train_set)
+
+
+# Write model hyper-parameters
+def write_params(model, path):
+  with open(path, 'w') as file:    
+    for stage in model.stages:
+          params = stage.extractParamMap()
+          for k in params:
+              file.write(f'{k.name}: {params[k]}\n')
+
+write_params(model, workdir + f'data/brf_{i}/params')
+
+
[…]

### 2019-04-23 weather.py
Change risky_weather feature to be only snow

@@ -107,7 +107,7 @@ def get_station_weather_month(station_id, year, month):
     # available at all, so it should not be considered as null risky_weater
     df['risky_weather'] = np.nan
     df.loc[df['Temp (°C)'].notna(), 'risky_weather'] = \
-        df['Weather'].astype(str).str.contains('snow|ice|freezing',
+        df['Weather'].astype(str).str.contains('snow',
                                                regex=True,
                                                flags=re.IGNORECASE)
 

### 2019-04-23 mains/main_negative_samples.py
Fix main_negative_samples.py

@@ -6,6 +6,6 @@ from workdir import workdir
 spark = init_spark()
 neg_samples = \
     get_negative_samples(spark,
-                         save_to=workdir + 'data/negative-sample-new.parquet',
+                         save_to='data/negative-sample-new.parquet',
                          sample_ratio=1e-3)
 print(neg_samples.count())

### 2019-04-23 random_forest.py
Change get_feature_importances: Sort feature importances

@@ -183,5 +183,5 @@ def get_feature_importances(model):
     feature_importances = pd.DataFrame(model.featureImportances.toArray())
     feature_importances.index = features_col
     feature_importances.columns = ["Feature importances"]
-    feature_importances.sort_values(['Feature importances'], ascending=False)
+    feature_importances = feature_importances.sort_values(['Feature importances'], ascending=False)
     return feature_importances

### 2019-04-23 preprocess.py
Fix sample_id column, now unique

@@ -432,6 +432,9 @@ def get_dataset_df(spark, pos_samples, neg_samples):
     pos_samples = pos_samples.withColumn('label', lit(1.0))
     neg_samples = neg_samples.withColumn('label', lit(0.0))
 
+    # Make sure sample_id is unique
+    neg_samples = neg_samples.withColumn('sample_id', -1 * col('sample_id'))
+
     pos_samples = pos_samples.select(*neg_samples.columns)
 
     train_pos, train_neg, test_pos, test_neg = \

### 2019-04-24 evaluate.py
Fix code style

@@ -19,9 +19,9 @@ def evaluate_binary_classifier(predictions):
                                       metricName="areaUnderROC")
     area_under_ROC = ROC_evaluator.evaluate(predictions)
     acc_evaluator = \
-    MulticlassClassificationEvaluator(predictionCol='prediction',
-                                      labelCol='label',
-                                      metricName='accuracy')
+        MulticlassClassificationEvaluator(predictionCol='prediction',
+                                          labelCol='label',
+                                          metricName='accuracy')
     acc_score = acc_evaluator.evaluate(predictions)
 
     print(f"Area Under PR = {area_under_PR}")

### 2019-04-24 random_forest.py

@@ -183,5 +183,7 @@ def get_feature_importances(model):
     feature_importances = pd.DataFrame(model.featureImportances.toArray())
     feature_importances.index = features_col
     feature_importances.columns = ["Feature importances"]
-    feature_importances = feature_importances.sort_values(['Feature importances'], ascending=False)
+    feature_importances = \
+        feature_importances.sort_values(['Feature importances'],
+                                        ascending=False)
     return feature_importances
